### 2016-10-23 README.tex
Refactored and added part of counter section.

\documentclass{tex/mjw}
\title{Enforced Invariant Confluence}
\author{Michael Whittaker}

\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{paralist}
\usepackage{tex/python}
\usepackage{tikz}

\newcommand{\figref}[1]{Figure \ref{fig:#1}}
\newcommand{\defeq}{\stackrel{\mathclap{\normalfont\mbox{\scriptsize def}}}{=}}
\newcommand{\dom}[1]{\text{dom}}

\newcommand{\iinvariant}{\mathcal{I}}
\newcommand{\iconfluent}{$\iinvariant$-confluent}
\newcommand{\iconfluence}{$\iinvariant$-confluence}

\begin{document}
\maketitle

\input{abstract.tex}
\input{vision.tex}
\input{counters.tex}

\bibliography{tex/bib}
\bibliographystyle{plain}
\end{document}