### 2016-10-24 doc/abstract.tex
Moved tex into doc directory.

\section{Abstract}
\emph{Strong consistency} allows programmers to ignore many of the complexities
of a distributed system, treating it as if it were running on a single machine.
However, enforcing strong consistency requires \emph{coordination}, and
coordination leads to unavailability \cite{gilbert2002brewer} (at worst) or
increased latency (at best). However, coordination cannot always be avoided.
Certain application invariants require coordination in order to be globally
enforced. Bailis et al. developed the notion of invariant confluence
(\iconfluence{}) as a necessary and sufficient condition for coordination
freedom \cite{bailis2014coordination}. Intuitively, a set of transactions is
\iconfluent{} with respect to some invariant $I$ if merging database states
that satisfy $I$ always results in a database state that also satisfies $I$.
Bailis characterized many common invariants (e.g. uniqueness constraints,
foreign key constraints), showing which could be maintained without
coordination. However, this characterization required hand-written proofs. This
research project aims to expand on Bailis' research by providing a set of
CRDTs \cite{shapiro2011comprehensive, shapiro2011conflict}, a language of
invariants, and a language of transactions that facilitates the ability for
\iconfluence{} to be determined algorithmically. An \iconfluence{} decision
procedure would allow programmers to guarantee that their programs can execute
without coordination.