### 2017-01-24 doc/paper/composition.tex
Divided paper into more sections.

\section{\iconfluence{}, Compositionally}\label{sec:composition}