### 2017-01-25 doc/paper/alie.tex
Revisions.

\section{\iconfluence{}, For Real This Time}\label{sec:alie}