### 2017-01-25 doc/paper/related.tex

\section{Related Work}\label{sec:related}
\todo{Cite and summarize \cite{balegas2015putting}, \cite{roy2015homeostasis},
\cite{li2012making}, \cite{li2014automating}, and \cite{gotsman2016cause}.}