### 2018-02-18 doc/paper_redux/iconfluence.tex
Added operation based approach to tex.

\section{\Iconfluence{}}
In this section, we present six different ways to think about \Iconfluence{}:
three state-based approaches and three operation-based approaches.

\subsection{State-Based}
\begin{definition}
  A \defword{distributed state-based object} is a triple $(S, s_0, \join)$
  where $S$ is a set of states, $s_0 \in S$ is a designated start state, and
  $\join: S \times S \to S$ is a binary merge operator that we call join.
\end{definition}

The notion of state-based objects is taken from~\cite{shapiro2011conflict}.

\begin{definition}
  A \defword{state-based transaction} $t: S \to S$ is a function that maps one
  state to another.
\end{definition}

\begin{definition}
  An \defword{invariant} $I: S \to \set{\text{true}, \text{false}}$ is a
  predicate on states.
\end{definition}

\begin{example}
  $(\nats, 0, +)$ is a distributed object, $t(x) = 2x$ is a transaction, and
  $I(x) = x \geq 0$ is an invariant.
\end{example}

There are three ways to think about state-based \Iconfluence{}: a process-based
approach, a graph-based approach, and an expression-based approach. All three
approaches are illustrated in \figref{statebasedmodels} and are described in
more detail below.

\input{statebased_iconfluence.fig}

\subsubsection{Process-Based}
The process-based model is most similar to the process model described by
Shapiro et al.\ in~\cite{shapiro2011conflict}. We are given a set
$\seq{p}{1}{n}$ of $n$ processors each of which begins with state $s_0$.
Processors can execute a transaction $t \in T$ to transition from state $s$ to
state $t(s)$, or they can send their state $s$ to another processor with state
$s'$ causing the other processor to transition from state $s'$ to state $s
\join s'$.

Consider any two processors $p_i$ and $p_j$ with states $s_i^0, \ldots, s_i^n$
and $s_j^0, \ldots, s_j^m$ such that every state $s_i^0, \ldots, s_i^n, s_j^0,
\ldots, s_j^m$ satisfies $I$. $T$ is invariant-confluent with respect to $I$,
abbreviated \defword{\Iconfluent{}}, if $s_i^n \join s_j^m$ is guaranteed to
satisfy $I$.

\subsubsection{Graph-Based}
The graph-based approach is most similar to the model used by Bailis et al.\
in~\cite{bailis2014coordination}. We are given a directed acyclic graph in
which vertices can have an arbitrary number of children but at most two
parents. The vertices are labelled with states, and the edges are either
labelled with transactions or are unlabelled if they correspond to a join. The
graph is really just an alternate way of representing an execution in the
process-based model, but with duplicate states collapsed into a single vertex.
As with the process-based model, we say that $T$ is \defword{\Iconfluent{}} if
the join of any two invariant satisfying states with invariant satisfying
ancestors is guaranteed to satisfy the invariant.

\subsubsection{Expression-Based}
The expression-based approach deals with expressions built from $s_0$,
transactions in $T$, and $\join$. That is, expressions $e$ are built from the
grammar
\[
  e ::= s_0 \mid t(e) \mid e_1 \join e_2
\]
where $t$ corresponds to a transaction in $T$. We can evaluate an expression,
denoted $eval(e)$, in the obvious way:
\begin{mathpar}
  eval(s_0) \defeq s_0

  eval(t(e)) \defeq t(eval(e))

  eval(e_1 \join e_2) \defeq eval(e_1) \join eval(e_2)
\end{mathpar}

\begin{definition}
  An expression $e$ \defword{satisfies $I$}, denoted $I(e)$, if $I(eval(e))$.
\end{definition}

\begin{definition}
  An expression \defword{recursively satisfies $I$}, denoted $\Irec{e}$, if $e$
  and all of $e$'s children satisfy $I$. That is,
  \begin{mathpar}
    \Irec{s_0} \defeq I(s_0)

    \Irec{t(e)} \defeq I(t(e)) \land \Irec{e}

    \Irec{e_1 \join e_2}
    \defeq I(e_1 \join e_2) \land \Irec{e_1} \land \Irec{e_2}
  \end{mathpar}
\end{definition}

\begin{definition}
  $T$ is \defword{\Iconfluent}, if for all expressions $e_1$ and $e_2$,
  $\Irec{e_1}$ and $\Irec{e_2}$ implies $I(e_1 \join e_2)$.
\end{definition}

\subsection{Operation-Based}
\begin{definition}
  A \defword{distributed operation-based object} is a pair $(S, s_0)$
  where $S$ is a set of states and $s_0 \in S$ is a designated start state.
\end{definition}

The notion of operation-based objects is also taken
from~\cite{shapiro2011conflict}. Note that we do not have a join function like
we did with state-based objects.

\begin{definition}
  An \defword{operation-based transaction} $t: S \to (S \to S)$ is a function
  that maps a state to a \defword{shadow transaction} $t(s): S \to S$. The term
  shadow transaction is taken from~\cite{li2014automating}.
\end{definition}

The definition of invariants remains unchanged.

\begin{example}
  $(\nats, 0)$ is a distributed object. Let $f_k(x) = x + k$. $t(x) = f_x$ is a
  transaction. $I(x) = x \geq 0$ is an invariant.
\end{example}

As with state-based objects, there are three ways to think about
operation-based \Iconfluence{}: a process-based approach, a graph-based
approach, and an expression-based approach. All three approaches are
illustrated in \figref{opbasedmodels} and are described in more detail below.

\input{opbased_iconfluence.fig}

\subsubsection{Process-Based}
As with the state-based approach, the process-based model is most similar to
the process model described by Shapiro et al.\ in~\cite{shapiro2011conflict}.
We are given a set $\seq{p}{1}{n}$ of $n$ processors each of which begins with
state $s_0$. Repeatedly, a processor executes a transaction $t$ on its current
state $s$ and then transitions from state $s$ to state $t(s)(s)$. It also
broadcasts $t(s)$ to all other processes exactly once. When a processor with
state $s'$ receives $t(s)$, it transitions from state $s'$ to state $t(s)(s')$.

Consider any two processors $p_i$ and $p_j$ with states $s_i^0, \ldots, s_i^n$
and $s_j^0, \ldots, s_j^m$ such that every state $s_i^0, \ldots, s_i^n, s_j^0,
\ldots, s_j^m$ satisfies $I$. $T$ is invariant-confluent with respect to $I$,
abbreviated \defword{\Iconfluent{}}, if for all $t \in T$, $I(t(s_i^n)(s_i^n))$
implies $I(t(s_i^n)(s_j^m))$.

\subsubsection{Graph-Based}
In the graph-based model, we are given a directed acyclic graph in which each
vertex is associated with a state $s$ and each edge is labelled with a shadow
operation $t(s)$ where $s$ is associated with some vertex in the graph.
%
Consider two invariant-satisfying states $s$ and $s'$ with invariant satisfying
ancestors. $T$ is \defword{\Iconfluent{}} if for all $t$, $I(t(s)(s))$ implies
$I(t(s)(s'))$.

\subsubsection{Expression-Based}
The expression-based approach deals with expressions built from $s_0$ and
transactions in $T$. That is, expressions $e$ are built from the grammar
\[
  e ::= s_0 \mid t(e_1)(e_2)
\]
where $t$ corresponds to a transaction in $T$. We can evaluate an expression,
denoted $eval(e)$, in the obvious way:
\begin{mathpar}
  eval(s_0) \defeq s_0

  eval(t(e_1)(e_2)) \defeq t(eval(e_1))(eval(e_2))
\end{mathpar}

\begin{definition}
  An expression $e$ \defword{satisfies $I$}, denoted $I(e)$, if $I(eval(e))$.
\end{definition}

\begin{definition}
  An expression \defword{recursively satisfies $I$}, denoted $\Irec{e}$, if $e$
  and all of $e$'s children satisfy $I$. That is,
  \begin{mathpar}
    \Irec{s_0} \defeq I(s_0)

    \Irec{t(e_1)(e_2)} \defeq I(t(e_1)(e_2)) \land \Irec{e_1} \land \Irec{e_2}
  \end{mathpar}
\end{definition}

\begin{definition}
  $T$ is \defword{\Iconfluent}, if for all expressions $e_1$ and $e_2$,
  $\Irec{e_1}$, $\Irec{e_2}$, and $I(t(e_1)(e_1))$ implies $I(t(e_1)(e_2))$.
\end{definition}