### 2018-02-20 doc/paper_redux/sufficient_conditions.tex
Added a slightly less silly example.

\section{Sufficient Conditions}
In this section, we present sufficient conditions for state-based and
operation-based invariant confluence. It turns out that these conditions are
sufficient but not necessary. However, if we relax the assumption of a fixed
start state $s_0$, then they are both sufficient and necessary.

\begin{claim}\clmlabel{statebasedsufficient}
  If for all states $s_1$ and $s_2$, $I(s_1)$ and $I(s_2)$ implies $I(s_1 \join
  s_2)$, then $T$ is \Iconfluent{}. In other words, if invariant satisfying
  states are closed under join, then $T$ is \Iconfluent.
\end{claim}
\begin{proof}
  Let $e_1, e_2$ be expressions such that $\Irec{e_1}$ and $\Irec{e_2}$. Then,
  $I(eval(e_1))$ and $I(eval(e_2))$, so $I(eval(e_1) \join eval(e_2))$ which
  implies $I(e_1 \join e_2)$.
\end{proof}

\begin{claim}\clmlabel{opbasedsufficient}
  If for all transactions $t \in T$ and for all states $s_1$ and $s_2$,
  $I(s_1)$ and $I(s_2)$ and $I(t(s_1)(s_1))$ implies $I(t(s_1)(s_2))$, then $T$
  is \Iconfluent{}.
\end{claim}
\begin{proof}
  Let $t \in T$ and $e_1, e_2$ be expressions such that $\Irec{e_1}$,
  $\Irec{e_2}$, and $I(t(e_1)(e_2))$. Let $s_1 = eval(e_1)$ and $s_2 =
  eval(s_2)$. Then, $I(s_1)$, $I(s_2)$, and $I(t(s_1)(s_1))$, so
  $I(t(s_1)(s_2))$. This implies $I(t(e_1)(e_2))$.
\end{proof}

While the conditions in \clmref{statebasedsufficient} and
\clmref{opbasedsufficient} are sufficient, they are not necessary.

\begin{claim}\clmlabel{statebasednotnecessary}
  If $T$ is \Iconfluent, it may not satisfy the condition in
  \clmref{statebasedsufficient}.
\end{claim}
\begin{proof}
  Consider the state based object $(S, s_0, \join)$ where $S$ is the set of all
  finite non-empty sets of integers, $s_0 = \set{0}$, and $\join = \cup$. We
  have a single transaction $t(x) = x \cup \set{\max(x) + 2}$ which retrieves
  the largest element from the set (which is guaranteed to exist because it is
  finite and non-empty), adds two to it, and then adds it into the set. $T =
  \set{t}$. Let $I$ be the invariant that all elements of $S$ are either odd or
  even.

  Note that $T$ is \Iconfluent{}. Beginning with our initial state, we can
  derive states $\set{0}, \set{0, 2}, \set{0, 2, 4}, \ldots$ all of which
  satisfy the invariant. Moreover, merging any such states satisfies the
  invariant.
  %
  However, $T$ doesn't satisfy the condition in \clmref{statebasedsufficient}.
  $\set{0}$ and $\set{1}$ satisfy the invariant, but $\set{0} \cup \set{1}$
  does not.
\end{proof}

\begin{claim}\clmlabel{opbasednotnecessary}
  If $T$ is \Iconfluent, it may not satisfy the condition in
  \clmref{opbasedsufficient}.
\end{claim}
\begin{proof}
  We play the same trick as \clmref{statebasednotnecessary}. Let $S$, $s_0$,
  and $I$ be as in \clmref{statebasednotnecessary}. Consider a single
  transaction $t(x) = u_{\max(x) + 2}$ where $u_k(x) = x \cup \set{k}$. Let $T
  = \set{t}$.

  $T$ is \Iconfluent{}. Beginning with our initial state, we can derive any
  finite set of even integers containing $0$. All these states satisfy the
  invariant. Letting $s, s'$ be any of these states, $I(t(s)(s'))$.
  %
  However, $T$ doesn't satisfy the condition in \clmref{opbasedsufficient}.
  $\set{0}$ and $\set{1}$ satisfy the invariant, and so does
  $t(\set{0})(\set{0})$, but $t(\set{0})(\set{1}) = \set{1, 2}$ does not.
\end{proof}

\todo{Think of more compelling examples.}

The conditions in \clmref{statebasedsufficient} and \clmref{opbasedsufficient}
are not necessary. Intuitively, this makes quite a bit of sense. The condition
in \clmref{statebasedsufficient} doesn't depend on $T$ and neither condition
depends on the initial state $s_0$. However, if we relax our assumption of a
fixed start state $s_0$ and instead allow replicas to begin with an arbitrary
state, then both conditions are sufficient and necessary. Alternatively, if
every state in $S$ is reachable from $s_0$, then the conditions are sufficient
and necessary.

\todo{Prove these facts more carefully. It's pretty straightforward though.}

\todo{Write this up in full.}
Imagine we are building a system to group students together. We have a relation
of (student id, grade) and a map (group id -> set of student ids). We want to
enforce the invariant that every student in every group is in the same grade.
If we start off with an initially empty mapping, then there are conflicts. If
we start off with at least one student mapped into each group, then there are
not any conflicts.

% | Data Model                              | I           | SC  | T | IC  |
% | State based set with union              | phi(x)      | yes |   | yes |
% | State based two sets with union         | foreign key | yes |   | yes |
% | State-based increment-decrement counter | >           |