### 2018-02-23 doc/paper_redux/iconfluence.tex
Started iconfluence definition tweaks.

\section{\Iconfluence{}}
In this section, we present six different ways to think about \Iconfluence{}:
three state-based approaches and three operation-based approaches.

\subsection{State-Based}
\begin{definition}
  A \defword{distributed state-based object} is a triple $(S, s_0, \join)$
  where $S$ is a set of states, $s_0 \in S$ is a designated start state, and
  $\join: S \times S \to S$ is a binary merge operator that we call join.
\end{definition}

The notion of state-based objects is taken from~\cite{shapiro2011conflict}.
Note that $\join$ is an \emph{arbitrary} function. It does not necessarily have
to satisfy any special properties like associativity or commutativity, though
later we will see that often it does.

\begin{definition}
  A \defword{state-based transaction} $t: S \to S$ is a function that maps one
  state to another.
\end{definition}

\begin{definition}
  An \defword{invariant} $I \subseteq S$ is a subset of states. Notationally,
  we say $I(s)$ to mean $s \in I$ and $\lnot I(s)$ to mean $s \notin I$.
\end{definition}

\begin{example}
  $(\nats, 0, +)$ is a distributed object, $t(x) = 2x$ is a transaction, and
  $\setst{x \in \nats}{x \equiv 0 \mod 2}$ is an invariant. Here, $I(0)$ and
  $I(2)$ but $\lnot I(1)$ and $\lnot I(3)$.
\end{example}

There are three ways to think about state-based \Iconfluence{}: a process-based
approach, a graph-based approach, and an expression-based approach. All three
approaches are illustrated in \figref{statebasedmodels} and are described in
more detail below.

\input{statebased_iconfluence.fig}

\subsubsection{Process-Based}
The process-based model is most similar to the process model described by
Shapiro et al.\ in~\cite{shapiro2011conflict}. We are given a set
$\seq{p}{1}{n}$ of $n$ processors each of which begins with state $s_0$.  A
processor $p_i$ can perform one of two actions.

First, $p_i$ can execute a transaction $t \in T$ to transition from state $s$
to state $t(s)$, given that $I(t(s))$. If $\lnot I(t(s))$, then $p_i$ will not
execute $t$ (or it will execute $t$ but then abort it; think of it however you
like).

Second, $p_i$ can send its state $s_i$ to another processor $p_j$ with state
$s_j$ causing $p_j$ to transition from state $s_j$ to state $s_i \join s_j$ (or
$s_j \join s_i$; it doesn't really matter). Note that unlike executing a
transaction, $p_j$ must transition from $s_j$ to $s_i \join s_j$ even if $\lnot
I(s_i \join s_j)$.

$T$ is invariant-confluent with respect to $I$, abbreviated \Iconfluent{}, if
every reachable state (including $s_0$) satisfies the invariant. Note that this
definition of \Iconfluence{} is different from the original definition
in~\cite{bailis2014coordination}. Later, we will see that they are equivalent.

\subsubsection{Graph-Based}
The graph-based approach is most similar to the model used by Bailis et al.\
in~\cite{bailis2014coordination}. We are given a directed acyclic graph where
vertices are states, and edges are either labelled with transactions or are
unlabelled if they correspond to a join. The graph is really just an alternate
way of representing an execution in the process-based model, but with duplicate
states collapsed into a single vertex. As with the process-based model, we say
that $T$ is \Iconfluent{} if every vertex in the graph satisfies the invariant.

\subsubsection{Expression-Based}
The expression-based approach formalizes the process-based and graph-based
approach using expressions. Given a state based object $(S, s_0, \join)$ and a
set of transactions $T$, we consider expressions generated by the following
grammar:
\[
  e ::= s_0 \mid t(e) \mid e_1 \join e_2
\]
where $t$ corresponds to a transaction in $T$. We can evaluate an expression
$e$, written $eval(e)$, in the obvious way:
\begin{mathpar}
  eval(s_0) \defeq s_0

  eval(t(e)) \defeq t(eval(e))

  eval(e_1 \join e_2) \defeq eval(e_1) \join eval(e_2)
\end{mathpar}

\begin{definition}
  An expression $e$ \defword{satisfies $I$}, denoted $I(e)$, if $I(eval(e))$.
\end{definition}

\begin{definition}
  Intuitively, we say an expression is \defword{reachable} if it can be reached
  in an execution of the process-based model. Formally, we defined a predicate
  $\reachable{\cdot}$ as follows:

  \begin{mathpar}
    \inferrule{ }{\reachable{s_0}}

    \inferrule{\reachable{e} \\ I(t(e))}{\reachable{t(e})}

    \inferrule{\reachable{e_1} \\ \reachable{e_2}}{\reachable{e_1 \join e_2}}
  \end{mathpar}
\end{definition}

\begin{definition}
  $T$ is \defword{\Iconfluent{}} if $\setst{e}{\reachable{e}} \subseteq I$. In
  other words, $T$ is \Iconfluent{} if all reachable states satisfy the
  invariant.
\end{definition}

As we mentioned earlier, this definition of \Iconfluence{} is different than
the original definition presented in~\cite{bailis2014coordination} (though it
is the same as the definition in~\cite{gotsman2016cause}), but they are
(almost) equivalent. We introduce some definitions and then prove this.

\begin{definition}
  An expression \defword{recursively satisfies $I$}, denoted $\Irec{e}$, if $e$
  and all of $e$'s children satisfy $I$. That is,
  \begin{mathpar}
    \Irec{s_0} \defeq I(s_0)

    \Irec{t(e)} \defeq I(t(e)) \land \Irec{e}

    \Irec{e_1 \join e_2}
    \defeq I(e_1 \join e_2) \land \Irec{e_1} \land \Irec{e_2}
  \end{mathpar}
\end{definition}

Bailis et al.\ defined \Iconfluence{} to mean that expressions recursively
satisfying $I$ are closed under join. That is, $T$ is \Iconfluent{} if
\[
  \forall e_1, e_2.\; \Irec{e_1} \land \Irec{e_2} \implies I{e_1 \join e_2}
\]
If $\lnot I(s_0)$, then \Iconfluence holds vacuously which is a bit silly, so
Bailis et al.\ probably ought to have also added the condition $I(s_0)$ to
their definition of \Iconfluence{}. Doing so, the two definitions become
equivalent.

\begin{claim}\clmlabel{irec_implies_reachable}
  $\Irec{e} \implies \reachable{e}$
\end{claim}
\begin{elidableproof}
  We perform a structural induction on $e$.
  \begin{itemize}
    \item \textbf{Case $s_0$.}
      Axiomatically, $\reachable{s_0}$.

    \item \textbf{Case $t(e)$.}
      $\Irec{t(e)}$, so $I(t(e))$ and $\Irec{e}$. By the inductive hypothesis,
      $\reachable{e}$, so by the definition of $\reachable{\cdot}$,
      $\reachable{t(e)}$.

    \item \textbf{Case $e_1 \join e_2$.}
      $\Irec{t(e)}$, so $\Irec{e_1}$ and $\Irec{e_2}$. By the inductive
      hypothesis, $\reachable{e_1}$ and $\reachable{e_2}$, so by the definition
      of $\reachable{\cdot}$, $\reachable{e_1 \join e_2}$.
  \end{itemize}
\end{elidableproof}

\begin{claim}
  Consider a state based object $(S, s_0, \join)$, a set of transactions $T$,
  and an invariant $I$. The following two are equivalent:
  \begin{enumerate}[\quad(1)]
    \item
      $\setst{e}{\reachable{e}} \subseteq I$

    \item
      $I(s_0)$ and $\Irec{e_1} \land \Irec{e_2} \implies I{e_1 \join e_2}$
  \end{enumerate}
\end{claim}
\begin{elidableproof}
  First, we show that (1) implies (2). Axiomatically, $\reachable{s_0}$, so by
  (1), $I(s_0)$. Let $e_1$ and $e_2$ be arbitrary expressions such that
  $\Irec{e_1}$ and $\Irec{e_2}$. By \clmref{irec_implies_reachable},
  $\reachable{e_1}$ and $\reachable{e_2}$. Thus, $\reachable{e_1 \join e_2}$,
  so by (1), $I(e_1 \join e_2)$.

  Next, we show that (2) implies (1). We prove by structural induction that for
  all $e$, $\reachable{e} \implies \Irec{e}$. From this, (1) is immediate.
  \begin{itemize}
    \item \textbf{Case $s_0$.}
      $I(s_0)$ by (2), so $\Irec{s_0}$

    \item \textbf{Case $t(e)$.}
      $\reachable{t(e)}$, so $\reachable{e}$ and $I(t(e))$. By the inductive
      hypothesis, $\Irec{e}$, so by the definition of $\Irec{\cdot}$,
      $\Irec{t(e)}$.

    \item \textbf{Case $e_1 \join e_2$.}
      $\reachable{e_1 \join e_2}$, so $\reachable{e_1}$ and $\reachable{e_2}$.
      By the inductive hypothesis, $\Irec{e_1}$ and $\Irec{e_2}$. By $(2)$,
      $I(e_1 \join e_2)$. Thus, $\Irec{e_1 \join e_2}$.
  \end{itemize}
\end{elidableproof}

\subsection{Operation-Based}
\begin{definition}
  A \defword{distributed operation-based object} is a pair $(S, s_0)$
  where $S$ is a set of states and $s_0 \in S$ is a designated start state.
\end{definition}

The notion of operation-based objects is also taken
from~\cite{shapiro2011conflict}. Note that we do not have a join function like
we did with state-based objects.

\begin{definition}
  An \defword{operation-based transaction} $t: S \to (S \to S)$ is a function
  that maps a state to a \defword{shadow transaction} $t(s): S \to S$. The term
  shadow transaction is taken from~\cite{li2014automating}.
\end{definition}

The definition of invariants remains unchanged.

\begin{example}
  $(\nats, 0)$ is a distributed object. Let $f_k(x) = x + k$. $t(x) = f_x$ is a
  transaction. $I(x) = x \geq 0$ is an invariant.
\end{example}

As with state-based objects, there are three ways to think about
operation-based \Iconfluence{}: a process-based approach, a graph-based
approach, and an expression-based approach. All three approaches are
illustrated in \figref{opbasedmodels} and are described in more detail below.

\input{opbased_iconfluence.fig}

\subsubsection{Process-Based}
As with the state-based approach, the process-based model is most similar to
the process model described by Shapiro et al.\ in~\cite{shapiro2011conflict}.
We are given a set $\seq{p}{1}{n}$ of $n$ processors each of which begins with
state $s_0$. Repeatedly, a processor executes a transaction $t$ on its current
state $s$ and then transitions from state $s$ to state $t(s)(s)$. It also
broadcasts $t(s)$ to all other processes exactly once. When a processor with
state $s'$ receives $t(s)$, it transitions from state $s'$ to state $t(s)(s')$.

Consider any two processors $p_i$ and $p_j$ with states $s_i^0, \ldots, s_i^n$
and $s_j^0, \ldots, s_j^m$ such that every state $s_i^0, \ldots, s_i^n, s_j^0,
\ldots, s_j^m$ satisfies $I$. $T$ is invariant-confluent with respect to $I$,
abbreviated \defword{\Iconfluent{}}, if for all $t \in T$, $I(t(s_i^n)(s_i^n))$
implies $I(t(s_i^n)(s_j^m))$.

\subsubsection{Graph-Based}
In the graph-based model, we are given a directed acyclic graph in which each
vertex is associated with a state $s$ and each edge is labelled with a shadow
operation $t(s)$ where $s$ is associated with some vertex in the graph.
%
Consider two invariant-satisfying states $s$ and $s'$ with invariant satisfying
ancestors. $T$ is \defword{\Iconfluent{}} if for all $t$, $I(t(s)(s))$ implies
$I(t(s)(s'))$.

\subsubsection{Expression-Based}
The expression-based approach deals with expressions built from $s_0$ and
transactions in $T$. That is, expressions $e$ are built from the grammar
\[
  e ::= s_0 \mid t(e_1)(e_2)
\]
where $t$ corresponds to a transaction in $T$. We can evaluate an expression,
denoted $eval(e)$, in the obvious way:
\begin{mathpar}
  eval(s_0) \defeq s_0

  eval(t(e_1)(e_2)) \defeq t(eval(e_1))(eval(e_2))
\end{mathpar}

\begin{definition}
  An expression $e$ \defword{satisfies $I$}, denoted $I(e)$, if $I(eval(e))$.
\end{definition}

\begin{definition}
  An expression \defword{recursively satisfies $I$}, denoted $\Irec{e}$, if $e$
  and all of $e$'s children satisfy $I$. That is,
  \begin{mathpar}
    \Irec{s_0} \defeq I(s_0)

    \Irec{t(e_1)(e_2)} \defeq I(t(e_1)(e_2)) \land \Irec{e_1} \land \Irec{e_2}
  \end{mathpar}
\end{definition}

\begin{definition}
  $T$ is \defword{\Iconfluent}, if for all expressions $e_1$ and $e_2$,
  $\Irec{e_1}$, $\Irec{e_2}$, and $I(t(e_1)(e_1))$ implies $I(t(e_1)(e_2))$.
\end{definition}

\todo{Define ``reachable'' states and describe that \Iconfluence{} is
equivalent to saying that reachable states are a subset of invariant satisfying
states.}
\todo{Maybe define Iconfluence to say that the initial state satisfies the
invariant?}