### 2018-02-26 doc/paper_redux/interactive_decision_procedure.tex
Show labels.

\section{Interactive \Iconfluence{} Decision Procedure}
Let us revisit \exampleref{statebased_not_necessary}

% \clmref{sufficientandnecessary} arms us with some potent understanding of
% \Iconfluence! Imagine we have an \Iclosed{} decision procedure that informs us
% that our state-based object from \exampleref{statebasednotnecessary} is not
% \Iclosed{}. Without \clmref{sufficientandnecessary}, we'd be a bit stuck. We'd
% have to seek out alternative means to determine whether or not our transactions
% were \Iconfluent{}. With the intuition provided by
% \clmref{sufficientandnecessary}, we can instead restrict our invariant to $I =
% \setst{(x, y)}{xy \leq 0 \land x \geq 0 \land y \leq 0}$. Now, the \Iclosed{}
% decision procedure informs us that we are indeed \Iclosed{}, so we know we're
% \Iconfluent{}.
%