### 2018-02-26 doc/paper_redux/interactive_decision_procedure.tex
Added interactive decision procedure.

\section{Interactive \Iconfluence{} Decision Procedure}
Imagine we have an \Iclosed{} decision procedure, and let's revisit
\exampleref{StateBasedNotNecessary}. We have a state-based object $O$, a set of
transactions $T$, and an invariant $I$. We want to know if $T$ is
\Iconfluent{}. We can hand $O$, $T$, and $I$ to the \Iclosed{} decision
procedure to see if $O$ is \Iclosed{}. If the decision procedure says yes, then
by \clmref{StateBasedSufficient}, we know that $T$ is \Iconfluent{}.

However, in our example, $O$ is not \Iclosed{}, so the decision procedure will
say no. \clmref{SufficientAndNecessary} tells us that this means one of two
things. Either (1) $T$ is not \Iconfluent{} or (2) $T$ is \Iconfluent{}, but
there are some invariant satisfying states that are not reachable. If we think
harder about the set of reachable states, we may realize that all reachable
states have a non-negative value of $x$. Thus, we may modify our invariant to
$\setst{(x, y)}{xy \leq 0 \land x \geq 0}$. Now, $O$ is \Iclosed{}! If we hand
$O$, $T$, and $I$ to the decision procedure, it will say yes, and we can
conclude that $T$ is \Iconfluent{}.

\newcommand{\comment}[1]{\State \textcolor{flatdenim}{\texttt{//} #1}}
\begin{algorithm}
  \caption{Interactive \Iconfluence{} Decision Procedure}%
  \algolabel{InteractiveDecisionProcedure}
  \begin{algorithmic}
    \Function{IsIconfluent}{$O$, $T$, $I$}
      \State \Return \Call{Helper}{$O$, $T$, $I$, $\emptyset$, $\emptyset$}
    \EndFunction

    \State

    \comment{$R$ is a set of reachable states.}
    \comment{$NR$ is a set of non-reachable states.}
    \Function{Helper}{$O$, $T$, $I$, $R$, $NR$}
      \State closed, $s_1$, $s_2$ $\gets$
               \Call{IsIclosed}{$O$, $T$, $(I \cap R) - NR$}
      \If {closed $=$ unknown}
        \State \Return unknown
      \ElsIf {closed $=$ yes}
        \State \Return{yes}
      \Else
        \State Augment $R$ and $NR$ with random search and user input
        \State Make sure $s_1$ and $s_2$ are included in $R$ or $NR$
        \If{$s_1, s_2 \in R$}
          \State \Return no
        \Else
          \State \Return \Call{Helper}{$O$, $T$, $I$, $R$, $NR$}
        \EndIf
      \EndIf
    \EndFunction
  \end{algorithmic}
\end{algorithm}

This example motivates \algoref{InteractiveDecisionProcedure}, a general
purpose interactive \Iconfluence{} decision procedure in which a user
iteratively refines the invariants until the decision procedure terminates.
Given an object $O$, a set of transactions $T$, and an invariant $I$, we call
\textsc{IsIconfluent}$(O, T, I)$ to determine if $T$ is \Iconfluent{}.
\textsc{IsIconfluent} assumes access to an \Iclosed{} decision procedure
\textsc{IsIclosed}$(O, T, I)$ which returns whether $O$ is \Iconfluent{} (yes,
no, or unknown). If it returns no, it also returns a counterexample $s_1$ and
$s_2$.

\todo{%
  Describe \textsc{IsIconfluent}. Describe how we can implement
  \textsc{IsIclosed} with Z3 and how users can specify $R$ and $NR$ with
  invariants.
}