### 2018-03-04 doc/paper_redux/interactive_decision_procedure.tex

\section{Interactive Invariant-Confluence Decision Procedure}
Imagine we have an \Iclosed{} decision procedure, and let's revisit
\exampleref{StateBasedNotNecessary}. We have a state-based object $O$, a start
state $s_0$, set of transactions $T$, and an invariant $I$. We want to know if
$O$ is \sTIconfluent{}. We can hand $O$ and $I$ to the \Iclosed{} decision
procedure to see if $O$ is \Iclosed{}. If the decision procedure says yes and
we can conclude that $I(s_0)$, then by \clmref{StateBasedSufficient}, we know
that $O$ is \sTIconfluent{}.

However, in our example, $O$ is not \Iclosed{}, so the decision procedure will
say no. \clmref{SufficientAndNecessary} tells us that this means one of two
things. Either (1) $O$ is \emph{not} \sTIconfluent{} or (2) $O$ is
\sTIconfluent{}, but there are some invariant satisfying states that are not
reachable. If we think harder about the set of reachable states, we may realize
that all reachable states have a non-negative value of $x$. Thus, we may modify
our invariant from $\setst{(x, y)}{xy \leq 0}$ to $\setst{(x, y)}{xy \leq 0
\land x \geq 0}$. Now, $O$ is \Iclosed{}! If we hand $O$ and $I$ to the
decision procedure, it will say yes. Moreover, $I(s_0)$, so we can conclude
that $O$ is \sTIconfluent{}.

\newcommand{\comment}[1]{\State \textcolor{flatdenim}{\texttt{//} #1}}
\begin{algorithm}
  \caption{Interactive Invariant-Confluence Decision Procedure}%
  \algolabel{InteractiveDecisionProcedure}
  \begin{algorithmic}
    \Function{IsIconfluent}{$O$, $s_0$, $T$, $I$}
      \State
        \Return \Call{Helper}{$O$, $s_0$, $T$, $I$, $\set{s_0}$, $\emptyset$}
    \EndFunction

    \State

    \comment{$R$ is a set of $\sTIreachable{}$ states.}
    \comment{$NR$ is a set of $\sTIunreachable{}$ states.}
    \Function{Helper}{$O$, $s_0$, $T$, $I$, $R$, $NR$}
      \If {$\lnot(I - NR)(s_0)$}
        \State \Return no
      \EndIf
      \State closed, $s_1$, $s_2$ $\gets$ \Call{IsIclosed}{$O$, $I - NR$}
      \If {closed $=$ unknown}
        \State \Return unknown
      \ElsIf {closed $=$ yes}
        \State \Return{yes}
      \Else
        \State Augment $R$ and $NR$ with random search and user input
        \State Make sure $s_1$ and $s_2$ are included in $R$ or $NR$
        \If{$s_1, s_2 \in R$}
          \State \Return no
        \Else
          \State \Return \Call{Helper}{$O$, $T$, $I$, $R$, $NR$}
        \EndIf
      \EndIf
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\todo{Rename IsIconfluent to IsSTIConfluent}
\newcommand{\IsIconfluent}{\textsc{IsIconfluent}}
\newcommand{\IsIclosed}{\textsc{IsIclosed}}
\newcommand{\Helper}{\textsc{Helper}}

This example motivates \algoref{InteractiveDecisionProcedure}, a general
purpose interactive invariant-confluence decision procedure in which a user
iteratively refines the invariant until the decision procedure terminates.
Given an object $O$, a start state $s_0$, a set of transactions $T$, and an
invariant $I$, we call \IsIconfluent$(O, s_0, T, I)$ to determine if $O$ is
\sTIconfluent{}. Now, we describe \algoref{InteractiveDecisionProcedure}. An
example of how to use \algoref{InteractiveDecisionProcedure} on
\exampleref{StateBasedNotNecessary} is given in
\figref{DecisionProcedureExample}.

\IsIconfluent{} assumes access to an \Iclosed{} decision procedure
\IsIclosed$(O, I)$ which returns whether $O$ is \Iclosed{} (yes, no, or
unknown). If it returns no, it also returns a counterexample $s_1$ and $s_2$
such that $I(s_1)$ and $I(s_2)$, but $\lnot I(s_1 \join s_2)$. Constructing an
\Iclosed{} decision procedure is hard (undecidable, I think), but we can do
pretty well using an SMT solver like Z3. There is also some existing research
on how to built smart \Iclosed{} decision procedures that we can
leverage~\cite{li2014automating}.

\IsIconfluent{} takes in $O$, $s_0$, $T$, and $I$. Formally, $I$ is set of
states (e.g. $\setst{(x, y)}{x \in \ints, y \in \ints, xy \leq 0}$), but an
actual implementation of \IsIconfluent{} would likely take in $I$ as a logical
formula (e.g. $xy \leq 0$). From now on, whenever we mention a (possibly
infinite) set of states, remember that we can always represent the set as a
finite formula.

\IsIconfluent{} calls a helper function \Helper{} which---in addition to $O$,
$s_0$, $T$, and $I$---takes as arguments a set $R$ of $\sTIreachable{}$ states
and a set $NR$ of $\sTIunreachable{}$ states. Like \IsIconfluent, \Helper$(O,
s_0, T, I, R, NR)$ returns whether $O$ is \sTIconfluent{} (assuming $R$ and
$NR$ are correct). Intuitively, $NR$ helps remove unreachable states from $I$
moving us closer to the case where $I \subseteq \setst{s}{\sTIreachable{s}}$.

\Helper{} first checks to see if $(I - NR)(s_0)$. If it is not, then \Helper{}
returns no. If $NR$ only contains unreachable states, then it doesn't contain
$s_0$, so $s_0 \in I - NR \iff s_0 \in I$. Thus, if $\lnot (I - NR)(s_0)$, then
$\lnot I(s_0)$, so $O$ is not \sTIconfluent.

Next, \Helper{} checks to see if $O$ is \iclosed{$(I - NR)$}. There are three
cases to consider.
\begin{itemize}
  \item
    If \IsIclosed{} gets stuck and doesn't know whether or not $O$ is
    \iclosed{$(I - NR)$}, then \Helper{} is stuck too and is forced to return
    unknown.

  \item
    If \IsIclosed{} determines that $O$ is \iclosed{$(I - NR)$}, then by
    \clmref{StateBasedSufficient}, $O$ is \sticonfluent{s_0}{T}{(I -
    NR)}, so
    $
      \setst{s}{\stireachable{s_0}{T}{(I-NR)}{s}}
        \subseteq I - NR
        \subseteq I
    $.
    If $NR$ only contains unreachable states, then
    $
      \setst{s}{\sTIreachable{s}} = \setst{s}{\stireachable{s_0}{T}{(I-NR)}{s}}
    $.
    Thus,
    $
      \setst{s}{\sTIreachable{s}} \subseteq I
    $.
    Thus, $O$ \sTIconfluent{}, so \Helper{} return yes.

  \item
    If \IsIclosed{} determines that $O$ is not \iclosed{$(I - NR)$}, then we
    have a counterexample $s_1$ and $s_2$. We want to determine whether $s_1$
    and $s_2$ are $\sTIreachable{}$ or $\sTIunreachable{}$. We can do so in two
    ways.

    First, we can randomly generate a set of reachable states and add them to
    $R$.

    Second, we can prompt the user to tell us whether or not the states are
    $\sTIreachable{}$ or $\sTIunreachable{}$. In addition to labelling $s_1$
    and $s_2$ as $\sTIreachable{}$ or not, the user can also refine $I$ by
    augmenting $NR$. Taking \exampleref{StateBasedNotNecessary} as an example,
    the user might look through the existing $\sTIreachable{}$ and
    $\sTIunreachable{}$ states and have an aha moment realizing that every
    point with a negative $x$-coordinate is $\sTIunreachable{}$. They can use
    this knowledge to add all points with negative $x$-coordinates to $NR$.

    After $s_1$ and $s_2$ have been labelled $\sTIreachable{}$ or not, we
    continue. If both $s_1$ and $s_2$ are $\sTIreachable{}$, then so is $s_1
    \join s_2$, but $\lnot I(s_1 \join s_2)$. Thus, $O$ is not \sTIconfluent{},
    so \Helper{} returns no. Otherwise, one of $s_1$ and $s_2$ is
    $\sTIunreachable{}$. Now, we recurse.
\end{itemize}

\input{figs/decision_procedure_example.fig}

% \algoref{InteractiveDecisionProcedure} is an \emph{interactive} decision
% procedure. It requires that a user classify states as reachable or unreachable.
% What happens if a user incorrectly classifies a state? Given a counterexample
% $s_1, s_2$, if a user incorrectly says that one of the unreachable points is
% reachable, then the $s_1, s_2 \in R$ check in \Helper{} might be
% erroneously met. In turn, \IsIconfluent{} might erroneously return no. This is
% inconvenient, but not the end of the world.
% Moreover, we can modify our decision procedure so that \Helper{}
% user provide a \emph{proof} of reachability whenever they label as: a series of transactions and joins that result in the