### 2018-03-05 doc/paper_redux/stratified_iconfluence.tex
Added section on stratified invariant confluence.

\section{Stratified Invariant-Confluence}
The interactive decision procedure in \algoref{InteractiveDecisionProcedure}
and the sufficient conditions in \clmref{NewStateBasedSufficient} give us two
ways to detect when an object $O$ is invariant-confluent. But what if it's not?
%
Let's revisit \exampleref{StateBasedNotNecessary} but with start state $s_0 =
(-2, 2)$. Now, $O$ is not \sTIconfluent{}. The set of $\sTIreachable{}$ points
(illustrated in \figref{FixingIconfluenceReachable}) is not a subset of $I$
(illustrated in \figref{FixingIconfluenceInvariant}).

This is unfortunate, but if we're determined to make $O$ \sTIconfluent{}? Well,
then we have to adjust the set of $\sTIreachable{}$ points so that they are a
subset of $I$. We can go about this in three ways:
\begin{itemize}
  \item \textbf{Adjust $s_0$.}
    $O$ is not \sticonfluent{(-2, 2)}{T}{I}, but it is \sticonfluent{(0,
    0)}{T}{I}. Thus, we can change our start state to become
    invariant-confluent. This is illustrated in \figref{FixingIconfluenceS}.

  \item \textbf{Adjust $T$.}
    $O$ is not \sticonfluent{s_0}{\set{t_{x+1}, t_{y-1}}}{I}, but it is
    \sticonfluent{s_0}{\set{t_{x+1}}}{I}. Thus, we can remove some transactions
    to become invariant-confluent. This is illustrated in
    \figref{FixingIconfluenceT}. Instead of outright forbidding some
    transactions, we can instead use locking to limit the concurrent execution
    of certain sets of transactions~\cite{balegas2015putting,
    gotsman2016cause}.

  \item \textbf{Adjust $I$.}
    $O$ is not \sticonfluent{s_0}{T}{\setst{(x, y)}{xy \leq 0}} but it is
    \sticonfluent{s_0}{T}{\setst{(x, y)}{x \leq 0, y \geq 0}}. Thus, we can
    remove some states from $I$ to become invariant-confluent. This is
    illustrated in \figref{FixingIconfluenceI}. We could also add some states
    to $I$ and become invariant-confluent, but it's likely that states are
    excluded from $I$ for good reason.
\end{itemize}

{\input{figs/fixing_iconfluence.fig}}

Unfortunately, none of these three adjustments are completely satisfactory.
There may only be one natural start state, and artificially making some
invariant-satisfying states unreachable---either by eliminating transactions or
restricting the invariant---is most likely a bit of a dirty hack. Moreover,
locking can sometimes be overly restrictive. For example, we can make the
example above invariant-confluent by requiring that both transactions $t_{x+1}$
and $t_{y-1}$ acquire a lock. However, once all processors have a state $(x,
y)$ where $x \geq 0$ and $y \leq 0$, the locks are no longer necessary.

Ideally, we would maintain the same start state, use locking to restrict
concurrently executing transactions as little as possible, and leave all
invariant-satisfying reachable states reachable. Can we achieve this?  Yes, at
least kind of. We'll do so using something we call \defword{stratified
invariant-confluence}. The main idea is that we'll partition the set of
invariant-satisfying states. Depending on the state we're in, we'll either
limit which transactions can execute concurrently, restrict the invariant, or
both.

Consider a state-based object $O = (S, \join)$, a start state $s_0$, a set of
transactions $T$, and an invariant $I$ deployed on $n$ processors $p_1, \ldots,
p_n$. Further assume that $O$ is a semilattice.
%
With stratified invariant-confluence, we construct an ordered sequence $P =
(I_1, T_1), \ldots, (I_n, T_n)$ of $n$ pairs $(I_i, T_i)$ with the following
properties:
\begin{itemize}
  \item
    $I_i \subseteq I$ for $1 \leq i \leq n$

  \item
    $T_i \subseteq T$ for $1 \leq i \leq n$

  \item
    $I = \cup_{i=1}^n I_i$

  \item
    For all $(I_i, T_i) \in P$ and all states $s \in I_i$, $O$ is
    \sticonfluent{s}{I_i}{T_i}.
\end{itemize}
All processors store a copy of $P$. Moreover, at any given point in time, a
processor $p_i$ will execute with one pair $(I_j, T_j) \in P$ \defword{active}.
We'll describe what this means in just a moment. Initially, every processor
$p_i$ is state $s_0$. $p_i$ finds the first pair $(I_j, T_j) \in P$ such that
$I_j(s_0)$ and makes this $(I_j, T_j)$ active. As usual, a processor can
perform one of two actions: execute a transaction or perform a merge.
\begin{itemize}
  \item
    Merging proceeds as usual. When a process $p_i$ with state $s_i$ receives a
    state $s_j$ from $p_j$, it transitions from $s_i$ to $s_i \join p_j$.

  \item
    Transaction processing is a bit different. Imagine a processor $p_i$ in
    state $s_i$ with pair $(I_j, T_j)$ active wants to execute transaction $t$.
    %
    If $t \in T_j$, then $p_i$ executes $t_i$ so long as doing so maintains
    $I_j$.
    %
    If $t \notin T_j$, then $p_i$ initiates a round of coordination among the
    processors. All processors adopt state state $s = s_1 \join \cdots \join
    s_n$. Then, all processors execute $t$. If $t(s)$ does not satisfy $I$,
    then the processors remain in state $s$ and continue normal processing.  If
    $t(s)$ does satisfy $I$, then the processors enter state $t(s)$ and choose
    pair $(I_k, T_k)$ to active where $(I_k, T_k)$ is the first pair in $P$
    where $I_k(t(s))$.
\end{itemize}

An example of stratified invariant-confluence for the example discussed above
is given in \figref{StratifiedIconfluenceExample}.

{\input{figs/stratified_iconfluence_example.fig}}

\todo{Describe why this scheme ensures ``Iconfluence''}.

\todo{%
  Give name to things like $P$. Call it something like an $(I, T)$-partition.
}
\todo{Can we modify the decision procedure to help build a stratification?}