### 2018-03-12 doc/paper_redux/evaluation.tex
Added thoughts on evaluation.

\section{Evaluation}
\begin{itemize}
  \item
    How good is the interactive decision procedure? How good are the new
    sufficient conditions? For this, we'll need a set of case studies to run
    our algorithms on. We can borrow case studies from related works.
  \item
    How fast is the interactive decision procedure? How fast are the new
    sufficient conditions? We can use the same case studies.
  \item
    How good is stratified invariant confluence? Find another set of case
    studies.
  \item
    How fast is Lucy for normal invariant confluence? Choose an application and
    measure the throughput or latency. There's not too many knobs to tune.
  \item
    How fast is Lucy for stratified invariant confluence? Choose an application
    and measure the throughput or latency. Vary the number of transactions that
    take us out of the region, the number of transactions outside the region,
    etc. Compare this to something like a primary-backup system where
    operations are all run one at a time.
\end{itemize}