### 2018-03-26 doc/ccn/diagrams.tex
Renaming various docs.

% Implication.
\tikzstyle{impl}=[-implies, double distance=3pt]

% Transaction.
\tikzstyle{txn}=[ultra thick, ->]

% Dashed transaction.
\tikzstyle{dtxn}=[ultra thick, ->, dashed]

% G transaction.
\tikzstyle{gtxn}=[ultra thick, ->, decorate, decoration={zigzag}]

% Dashed G transaction.
\tikzstyle{dgtxn}=[ultra thick, ->, densely dashed, decorate, decoration={zigzag}]

% Invariant node.
\newcommand{\inode}[3][]{
  \node[draw, shape=circle, minimum width=0.6cm] (#2) at (#3) {};
  \node[fill, shape=circle, #1] () at (#3) {};
}

% Dashed node.
\newcommand{\dnode}[3][]{
  \node[draw, dashed, shape=circle, minimum width=0.6cm] (#2) at (#3) {};
  \node[fill, shape=circle, #1] () at (#3) {};
}

% Normal node.
\newcommand{\nnode}[3][]{
  \node[fill, shape=circle, #1] (#2) at (#3) {};
}