### 2018-03-29 doc/paper/sections/invariant_closure.tex
Rough outline + some of iconfluence section.

\section{Invariant-Closure}
% - define invariant-closure and introduce it as a sufficient condition
% - cite existing work to show that people have used invariant closure to
%   reason about invariant-confluence
% - explain why invariant closure can fail and why its not sufficient using the
%   two-integers example
% - state main theorem