### 2018-03-29 doc/paper/sections/invariant_closure.tex
Finished invariant-closure section.

\section{Invariant-Closure}\seclabel{InvariantClosure}
Our ultimate goal is to write a program that can automatically decide whether a
given distributed object $O$ is \sTIconfluent{}. Such a program has to
automatically prove or disprove that every reachable state satisfies the
invariant. However, automatically reasoning about the possibly infinite set of
reachable states is challenging, especially because transactions and merge
functions can be complex and can be interleaved arbitrarily in an execution.
Due to this complexity, existing systems that aim to automatically decide
invariant-confluence instead focus on deciding a sufficient condition for
invariant-confluence---dubbed \defword{invariant-closure}---that is simpler to
reason about~\cite{li2012making, li2014automating}. In this section, we define
invariant-closure and study why the condition sufficient but not necessary.
Armed with this understanding, we present conditions under which it is both
sufficient and necessary.

We say an object $O = (S, \join)$ is \defword{invariant-closed with respect to
an invariant $I$}, abbreviated \defword{\Iclosed{}}, if invariant-satisfying
states are closed under merge. That is, for every state $s_1, s_2 \in S$, if
$I(s_1)$ and $I(s_2)$, then $I(s_1 \join s_2)$.

\begin{theorem}\thmlabel{IclosureImpliesIconfluence}
  Given an object $O = (S, \join)$, a start state $s_0 \in S$, a set of
  transactions $T$, and an invariant $I$, if $I(s_0)$ and if $O$ is \Iclosed{},
  then $O$ is \sTIconfluent{}.
\end{theorem}

\thmref{IclosureImpliesIconfluence} states that invariant-closure is
sufficient for invariant-confluence. Intuitively, our system model ensures that
transaction execution preserves the invariant, so if merging states also
preserves the invariant and if our start state satisfies the invariant, then
inductively it is impossible for us to reach a state which doesn't satisfy the
invariant.

This is good news because checking if an object is invariant-closed is more
straightforward than checking if it is invariant-confluent. Existing systems
typically use an SMT solver like Z3 to check if an object is
invariant-closed~\cite{de2008z3, balegas2015putting, gotsman2016cause}. If it
is, then by \thmref{IclosureImpliesIconfluence}, it is invariant-confluent.
Unfortunately, invariant-closure is \emph{not} necessary for
invariant-confluence, so if an object is not invariant-closed, these systems
cannot conclude that the object is not invariant-confluent. The reason why
invariant-closure is not necessary for invariant-confluence is best explained
through an example.

\begin{example}\examplelabel{Z2}
  Let $O = (\ints \times \ints, \join)$ consist of pairs $(x, y)$ of integers
  where $(x_1, y_1) \join (x_2, y_2) = (\max(x_1, x_2), \max(y_1, y_2))$. Our
  start state $s_0 \in \ints \times \ints$ is the point $(0, 0)$. Our set $T$
  of transactions consists of two transactions: $t_{x+1}((x, y)) = (x + 1, y)$
  which increments $x$ and $t_{y-1}((x, y)) = (x, y - 1)$ which decrements $y$.
  Our invariant $I = \setst{(x, y) \in \ints \times \ints}{xy \leq 0}$ consists
  of all points $(x, y)$ where the product of $x$ and $y$ is non-positive.

  The invariant and the set of reachable states are illustrated in \figref{Z2}
  in which we draw each state $(x, y)$ as a point in space. The invariant
  consists of the second and fourth quadrant, and the reachable states consist
  of the fourth quadrant. From this, it's immediate that the reachable states
  are a subset of the invariant, so $O$ is invariant-confluent. However,
  letting $s_1 = (-1, 1)$ and $s_2 = (1, -1)$, we see that $O$ is not
  invariant-closed. $I(s_1)$ and $I(s_2)$, but letting $s_3 = s_1 \join s_2 =
  (1, 1)$, we see $\lnot I(s_3)$.
\end{example}

{\input{figures/z2}}

In \exampleref{Z2}, $s_1$ and $s_2$ witness the fact that $O$ is not
invariant-closed, but $s_1$ is not reachable. This is not particular to
\exampleref{Z2}. In fact, it is fundamentally the reason why invariant-closure
is not equivalent to invariant-confluence. Invariant-confluence is, at its
core, a property of reachable states, but invariant-closure is completely
ignorant of reachability. As a result, invariant-satisfying yet unreachable
states like $s_1$ are the key hurdle preventing invariant-closure from being
equivalent to invariant-confluence. This is formalized by
\thmref{IclosureEquivalentIconfluence}

\begin{theorem}\thmlabel{IclosureEquivalentIconfluence}
  Consider an object $O = (S, \join)$, a start state $s_0 \in S$, a set of
  transactions $T$, and an invariant $I$. If the invariant is a subset of the
  reachable states (i.e. $I \subseteq \setst{s \in S}{\sTIreachable{s}}$), then
  \[
    \text{$I(s_0)$ and $O$ is \Iclosed{}} \iff \text{$O$ is \sTIconfluent{}}
  \]
\end{theorem}
\begin{proof}
  The forward direction of \thmref{IclosureEquivalentIconfluence} follows
  immediately from \thmref{IclosureImpliesIconfluence}. The backward direction
  holds because any two invariant satisfying states $s_1$ and $s_2$ must be
  reachable (by assumption), so their join $s_1 \join s_2$ is also reachable.
  And because $O$ is \sTIconfluent{}, all reachable points, including $s_1
  \join s_2$, satisfy the invariant.
\end{proof}

% - define invariant-closure and introduce it as a sufficient condition
% - cite existing work to show that people have used invariant closure to
%   reason about invariant-confluence
% - explain why invariant closure can fail and why its not sufficient using the
%   two-integers example
% - state main theorem