### 2018-03-29 doc/paper/sections/interactive_decision_procedure.tex
Progress on interactive decision procedure section.

\section{An Interactive Decision Procedure}\seclabel{InteractiveDecisionProcedure}
\newcommand{\IsInvConfluent}{\textsf{IsInvConfluent}}
\newcommand{\Helper}{\textsf{Helper}}

\newcommand{\algocomment}[1]{\State \textcolor{flatdenim}{\texttt{//} #1}}
\begin{algorithm}[t]
  \caption{Interactive Invariant-Confluence Decision Procedure}%
  \algolabel{InteractiveDecisionProcedure}
  \begin{algorithmic}
    \Function{IsInvConfluent}{$O$, $s_0$, $T$, $I$}
    \State
    \Return $I(s_0)$ and
    \Call{Helper}{$O$, $s_0$, $T$, $I$, $\set{s_0}$, $\emptyset$}
    \EndFunction

    \State

    \algocomment{$R$ is a set of $\sTIreachable{}$ states.}
    \algocomment{$NR$ is a set of $\sTIunreachable{}$ states.}
    \algocomment{$I(s_0)$ is a precondition.}
    \Function{Helper}{$O$, $s_0$, $T$, $I$, $R$, $NR$}
    \State closed, $s_1$, $s_2$ $\gets$ \Call{IsIclosed}{$O$, $I - NR$}
    \If {closed}
      \State \Return true
    \EndIf
    \State Augment $R$ and $NR$ with random search and user input
    \If{$s_1, s_2 \in R$}
    \Return false
    \Else\
    \Return \Call{Helper}{$O$, $s_0$, $T$, $I$, $R$, $NR$}
    \EndIf
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\thmref{IclosureEquivalentIconfluence} tells us that if all
invariant-satisfying points are reachable, then invariant-closure and
invariant-confluence are equivalent. In this section, we present the
interactive invariant-confluence decision procedure shown in
\algoref{InteractiveDecisionProcedure} which takes advantage of this result.

A user provides the algorithm with an object $O = (S, \join)$, a start state
$s_0$, a set of transactions $T$, and an invariant $I$\footnote{%
  Formally, $I$ is a (possibly infinite) set of states (e.g. $\setst{(x, y) \in
  \ints \times \ints}{xy \leq 0}$), but in practice, our implementation of
  \IsInvConfluent{} takes in $I$ as a finite logical formula (e.g. $xy \leq
  0$).
}. The user then interacts with
\algoref{InteractiveDecisionProcedure} to iteratively eliminate unreachable
states from the invariant. Meanwhile, the algorithm leverages an
invariant-closure decision procedure to either (a) conclude that $O$ is or is
not \sTIconfluent{} or (b) provide counterexamples to the user to help them
eliminate unreachable states. After all unreachable states have been eliminated
from the invariant, \thmref{IclosureEquivalentIconfluence} allows us to reduce
the problem of invariant-confluence directly to the problem of
invariant-closure, and the algorithm terminates.
%
We now describe \algoref{InteractiveDecisionProcedure} in detail. An example of
how to use \algoref{InteractiveDecisionProcedure} on \exampleref{Z2} is given
in \figref{DecisionProcedureExample}.

% This example motivates \algoref{InteractiveDecisionProcedure}, a general
% purpose interactive invariant-confluence decision procedure in which a user
% iteratively refines the invariant until the decision procedure terminates.
% Given an object $O$, a start state $s_0$, a set of transactions $T$, and an
% invariant $I$\footnote{%
%   Formally, $I$ is set of states (e.g. $\setst{(x, y)}{x \in \ints, y \in
%   \ints, xy \leq 0}$), but an actual implementation of \IsInvConfluent{} would
%   likely take in $I$ as a logical formula (e.g. $xy \leq 0$). From now on,
%   whenever we mention a (possibly infinite) set of states, remember that we can
%   always represent the set as a finite formula.
% }, we call \IsInvConfluent$(O, s_0, T, I)$ to determine if $O$ is
% \sTIconfluent{}. Now, we describe \algoref{InteractiveDecisionProcedure}. An
% example of how to use \algoref{InteractiveDecisionProcedure} on
% \exampleref{StateBasedNotNecessary} is given in
% \figref{DecisionProcedureExample}.
%
% \IsInvConfluent{} assumes access to two helper decision procedures:
% \begin{enumerate}
%   \item
%     A decision procedure to determine $I(s_0)$.
%
%   \item
%     An \Iclosed{} decision procedure \IsIclosed$(O, I)$ which returns a triple
%     (closed, $s_1$, $s_2$). closed is a boolean indicating whether $O$ is
%     \Iclosed{}. If closed is false, then $s_1$ and $s_2$ are a counterexample
%     such that $I(s_1)$ and $I(s_2)$, but $\lnot I(s_1 \join s_2)$.
% \end{enumerate}
%
% Constructing these two helper decision procedures is hard (undecidable, I
% think), but we can do pretty well using an SMT solver like Z3. There is also
% some existing research on how to built smart \Iclosed{} decision procedures
% that we can leverage~\cite{li2014automating}. In practice either decision an
% return unknown instead of true or false. In this case, \IsInvConfluent{} returns
% unknown.
%
% \IsInvConfluent{} first checks that $s_0$ satisfies $I$. If it doesn't, then
% \IsInvConfluent{} returns false. If it does then, \IsInvConfluent{} calls a
% helper function \Helper{} which---in addition to $O$, $s_0$, $T$, and
% $I$---takes as arguments a set $R$ of $\sTIreachable{}$ states and a set $NR$
% of $\sTIunreachable{}$ states. Like \IsInvConfluent, \Helper$(O, s_0, T, I, R,
% NR)$ returns whether $O$ is \sTIconfluent{} (assuming $R$ and $NR$ are
% correct). Intuitively, $NR$ helps remove unreachable states from $I$ moving us
% closer to the case where $I \subseteq \setst{s}{\sTIreachable{s}}$.
%
% First, \Helper{} checks to see if $O$ is \iclosed{(I - NR)}. There are two
% cases to consider.
% \begin{itemize}
%   \item
%     If \IsIclosed{} determines that $O$ is \iclosed{(I - NR)}, then by
%     \clmref{StateBasedSufficient}, $O$ is \sticonfluent{s_0}{T}{(I -
%     NR)}, so
%     $
%       \setst{s}{\stireachable{s_0}{T}{(I-NR)}{s}}
%         \subseteq I - NR
%         \subseteq I
%     $.
%     If $NR$ only contains unreachable states, then
%     $
%       \setst{s}{\sTIreachable{s}} = \setst{s}{\stireachable{s_0}{T}{(I-NR)}{s}}
%     $.
%     Thus,
%     $
%       \setst{s}{\sTIreachable{s}} \subseteq I
%     $.
%     Thus, $O$ \sTIconfluent{}, so \Helper{} return true.
%
%   \item
%     If \IsIclosed{} determines that $O$ is not \iclosed{(I - NR)}, then we
%     have a counterexample $s_1$ and $s_2$. We want to determine whether $s_1$
%     and $s_2$ are $\sTIreachable{}$ or $\sTIunreachable{}$. We can do so in two
%     ways.
%
%     First, we can randomly generate a set of reachable states and add them to
%     $R$.
%     %
%     Second, we can prompt the user to tell us whether or not the states are
%     $\sTIreachable{}$ or $\sTIunreachable{}$. In addition to labelling $s_1$
%     and $s_2$ as $\sTIreachable{}$ or not, the user can also refine $I$ by
%     augmenting $R$ and $NR$ arbitrarily (see \figref{DecisionProcedureExample}
%     for an example). In this step, we also make sure that $s_0 \notin NR$ since
%     we know that $s_0$ is $\sTIreachable{}$.
%
%     After $s_1$ and $s_2$ have been labelled $\sTIreachable{}$ or not, we
%     continue. If both $s_1$ and $s_2$ are $\sTIreachable{}$, then so is $s_1
%     \join s_2$, but $\lnot I(s_1 \join s_2)$. Thus, $O$ is not \sTIconfluent{},
%     so \Helper{} returns false. Otherwise, one of $s_1$ and $s_2$ is
%     $\sTIunreachable{}$, so we recurse.
% \end{itemize}
%
% \algoref{InteractiveDecisionProcedure} is an \emph{interactive} decision
% procedure. It requires that a user classify states as reachable or unreachable.
% What happens if a user incorrectly classifies a state? There's two possible
% errors that can be made:
% \begin{itemize}
%   \item \textbf{A user can label an unreachable state as reachable.}
%     In this case, the $s_1, s_2 \in R$ check in \Helper{} might be erroneously
%     met. In turn, \IsInvConfluent{} might erroneously return no. This is
%     inconvenient, but not the end of the world. Moreover, we can modify our
%     decision procedure so that \Helper{} requires that whenever a user labels a
%     state $s$ as $\sTIreachable{}$, they must also provide an $\sTIreachable{}$
%     expression $e$ where $eval(e) = s$.
%
%   \item \textbf{A user can label a reachable state as unreachable.}
%     In this case, \IsIclosed$(O, I - NR)$ might return true, even though $O$ is
%     not \sTIconfluent. Thus, a user might falsely believe their system to be
%     \sTIconfluent{} even though it isn't, and eventually their system might
%     enter a state which violates the invariant.
%
%     We can mitigate this in two ways. First, we can aggresively expand $R$. If
%     a user ever labels a state $s$ as unreachable, but $s \in R$, we can notify
%     the user of their mistake. Second, \Helper{} returns true when $O$ is
%     \iclosed{(I - NR)} for some $NR$, so $O$ is \sticonfluent{s_0}{T}{(I-R)}
%     (even though it might not be \sTIconfluent{}). Thus, when a user deploys
%     their system, they can deploy with the invariant $I - NR$ instead of $NR$.
%     If $NR$ is correctly populated, then deploying with $I - NR$ has no effect
%     on the system. If $NR$ is incorrectly populated, then some
%     $\sTIreachable{}$ states are artificially made unreachable, but the system
%     is still guaranteed to never enter a state which violates $I$.
% \end{itemize}
%

{\input{figures/decision_procedure_example}}

% - describe how main theorem gives us intution about interactive dp
% - present the algorithm
% - walk through an example
% - describe how the algorithm is robust to errors