### 2018-03-31 doc/paper/figures/merge_reducible.tex
Finished merge reduction section.

\begin{figure}[t]
  \centering
  \begin{tikzpicture}[scale=0.3]
    \draw[rounded corners] (0, 0) rectangle (16, 8);
    \draw[rounded corners] (1, 2.5) rectangle (11, 6.5);
    \draw[rounded corners] (6, 0.5) rectangle (15, 4.5);
    \draw[rounded corners] (7, 1.5) rectangle ++(7, 2);
    \node[inner sep=0pt, anchor=north] at (8, 7.5) {\small Invariant-Confluent};
    \node[inner sep=0pt, anchor=north] at (6, 6) {\small Invariant-Closed};
    \node[inner sep=0pt, anchor=south] at (10.5, 1.5) {\small Criteria (1) -- (4)};
    \node[inner sep=0pt, anchor=south] at (10.5, 0.5) {\small Merge-Reducible};
  \end{tikzpicture}
  \caption{%
    The relationship between invariant-closure, merge-reducibility, criteria
    (1) -- (4) from \thmref{LatticeProperty}, and invariant-confluence.
  }\figlabel{IclosureVsReducible}
\end{figure}