### 2018-04-05 doc/paper/sections/segmented_invariant_confluence.tex
Documented segmented invariant-confluence DP.

\section{Segmented Invariant-Confluence}\seclabel{SegmentedInvariantConfluence}
\newcommand{\IsIclosed}{\textsf{IsIclosed}}

If a distributed object is invariant-confluent, then the object can be
replicated without the need for any form of coordination to maintain the
object's invariant. But what if the object is \emph{not} invariant-confluent?
In this section, we present a generalization of invariant-confluence called
\defword{segmented invariant-confluence} that can be used to maintain the
invariants of non-invariant-confluent objects, requiring only a small amount of
coordination.

The main idea behind segmented invariant-confluence is to segment the state
space into a number of segments and restrict the set of allowable transactions
within each segment in such a way that the object is invariant-confluent
\emph{within each segment} (even though it may not be globally
invariant-confluent). Then, servers can run coordination-free within a segment
and need only coordinate when transitioning from one segment to another. We now
formalize segmented invariant-confluence, describe the system model we use to
replicate segmented invariant-confluent objects, and introduce a segmented
invariant-confluence decision procedure.

\subsection{Formalism}
Consider a distributed object $O = (S, \join)$, a start state $s_0 \in S$, a
set of transitions $T$, and an invariant $I$. A segmentation $\Sigma = (I_1,
T_1), \ldots, (I_n, T_n)$ is a sequence of $n$ segments $(I_i, T_i)$ where
every $T_i$ is a subset of $T$. $O$ is \defword{segmented invariant-confluent}
with respect to $s_0$, $T$, $I$, and $\Sigma$, abbreviated
\defword{\sTISconfluent}, if the following conditions hold:
\begin{itemize}
  \item
    The start state satisfies the invariant (i.e. $I(s_0)$).

  \item
    $I$ is covered by the invariants in $\Sigma$ (i.e. $I = \cup_{i=1}^n I_i$).
    Note that invariants in $\Sigma$ do \emph{not} have to be disjoint. Thus,
    they do not have to partition $I$; they just have to cover $I$.

  \item
    $O$ is invariant-confluent within each segment. That is, for every $(I_i,
    T_i) \in \Sigma$ and for every state $s \in I_i$, $O$ is
    \sticonfluent{s}{T_i}{I_i}.
\end{itemize}

{\input{figures/segmented_z2}}

\begin{example}\examplelabel{SegmentedZ2}
  Consider again the object $O = (\ints \times \ints, \join)$, transactions $T
  = \set{t_{x+1}, t_{y-1}}$, and invariant $I = \setst{(x, y)}{xy \leq 0}$ from
  \exampleref{Z2}, but now let the start state $s_0$ be $(-42, 42)$. $O$ is
  \emph{not} \sTISconfluent{} because the points $(0, 42)$ and $(42, 0)$ are
  reachable, and merging these points yields $(42, 42)$ which violates the
  invariant. However, $O$ is \sTISconfluent{} for $\Sigma = (I_1, T_1)$, $(I_2,
  T_2)$, $(I_3, T_3)$, $(I_4, T_4)$ where
  \begin{align*}
    I_1 &= \setst{(x, y)}{x < 0, y > 0}       &T_1 &= \set{t_{x+1}, t_{y-1}} \\
    I_2 &= \setst{(x, y)}{x \geq 0, y \leq 0} &T_2 &= \set{t_{x+1}, t_{y-1}} \\
    I_3 &= \setst{(x, y)}{x = 0}              &T_3 &= \set{t_{y-1}} \\
    I_4 &= \setst{(x, y)}{y = 0}              &T_4 &= \set{t_{x+1}}
  \end{align*}
  $\Sigma$ is illustrated in \figref{SegmentedZ2}. Clearly, $s_0$ satisfies the
  invariant, and $I_1, I_2, I_3, I_4$ cover $I$. Moreover, for every $(I_i,
  T_i) \in \Sigma$, we see that $O$ is \iclosed{I_i}, so $O$ is
  \sticonfluent{s}{T_i}{I_I} for every $s \in I_i$. Thus, $O$ is
  \sTISconfluent{}.
\end{example}

\subsection{System Model}
Now, we describe the system model used to replicate a segmented
invariant-confluent object without any coordination within a segment and with
only a small amount of coordination when transitioning between segments. As
before, we replicate an object $O$ across a set $p_1, \ldots, p_n$ of $n$
servers each of which manages a replica $s_i \in S$ of the object. Every server
begins with $s_0$, $T$, $I$, and $\Sigma$. Moreover, at any given point in
time, a server designates one of the segments in $\Sigma$ as its
\defword{active segment}.  Initially, every server chooses the first segment
$(I_i, T_i) \in \Sigma$ such that $I_i(s_0)$ to be its active segment. We'll
see momentarily the significance of the active segment.

As before, servers repeatedly perform one of two actions: execute a transaction
or merge states with another server. Merging states in the segmented
invariant-confluence system model is identical to merging states in the
invariant-confluence system model. A server $p_i$ sends its state $s_i$ to
another server $p_j$ which \emph{must} merge $s_i$ into its state $s_j$.
%
Transaction execution in the new system model, on the other hand, is a bit more
involved. Consider a server $s_i$ with active segment $(I_i, T_i)$. A client
can request that $p_i$ execute a transaction $t \in T$. We consider what
happens when $t \in T_i$ and $t \notin T_i$ separately.

If $t \notin T_i$, then $p_i$ initiates a round of global coordination to
execute $t$. During global coordination, every server temporarily stops
processing and transitions to state $s = s_1 \join \ldots \join s_n$, the join
of every server's state. Then, every server speculatively executes $t$
transitioning to state $t(s)$. If $t(s)$ violates the invariant (i.e.  $\lnot
I(t(s))$), then every server aborts $t$ and reverts to state $s$.  Then, $p_i$
replies to the client. If $t(s)$ satisfies the invariant (i.e. $I(t(s))$), then
every server commits $t$ and remains in state $t(s)$. Every server then chooses
the first segment $(I_i, T_i) \in \Sigma$ such that $I_i(t(s))$ to be the new
active segment. Note that such a segment is guaranteed to exist because the
segment invariants cover $I$. Moreover, $\Sigma$ is ordered, so every server
will deterministically pick the same active segment. In fact, an invariant of
the system model is that at any given point of normal processing, every server
has the same active segment.

$t \in T_i$, then $p_i$ executes $t$ immediately and without coordination. If
$t(s_i)$ satisfies the \emph{active} invariant (i.e. $I_i(t(s_i))$), then $p_i$
commits $t$, stays in state $t(s_0)$, and replies to the client. If $t(s_i)$
violates the invariant (i.e.  $i.e. \lnot I(t(s_i))$), then $p_i$ aborts $t$,
reverts to state $s_i$, and replies to the client. If $t(s_i)$ satisfies the
invariant but violates the active invariant (i.e. $I(t(s_i))$ but $\lnot
I_i(t(s_i))$), then $p_i$ reverts to state $s_i$ and initiates a round of
global coordination to execute $t$, as described in the previous paragraph.
Transaction execution is summarized in \algoref{TxnExecution}.

\begin{algorithm}[t]
  \caption{Transaction execution in segmented invariant-confluence system model}%
  \algolabel{TxnExecution}
  \begin{algorithmic}
    \If{$t \notin T_i$}
      \State Execute $t$ with global coordination
    \Else{}
      \If{$I_i(t(s_i))$} Commit $t$
      \ElsIf{$\lnot I(t(s_i))$} Abort $t$
      \Else{} Execute $t$ with global coordination
      \EndIf{}
    \EndIf{}
  \end{algorithmic}
\end{algorithm}

This system model guarantees that all replicas of a segmented
invariant-confluent object always satisfy the invariant. All servers begin in
the same initial state and with the same active segment. Thus, because $O$ is
invariant-confluent with respect to every segment, servers can execute
transactions within the active segment without any coordination and guarantee
that the invariant is never violated. Any operation that would violate the
assumptions of invariant-confluence system model (e.g.\ executing a transaction
that's not permitted in the active segment or executing a permitted transaction
that leads to a state outside the active segment) triggers a global
coordination. Globally coordinated transactions are only executed if they
maintain the invariant. Moreover, if a globally coordinated transaction leads
to another segment, the coordination ensures that all servers begin in the same
start state and with the same active replica, reestablishing the assumptions of
the invariant-confluence system model.

\subsection{Interactive Decision Procedure}
In order to determine whether or not an object $O$ is \sTISconfluent{} we have
to determine whether or not $O$ is invariant-confluent within each segment
$(I_i, T_i) \in \Sigma$. That is, we have determine if $O$ is
\sticonfluent{s}{T_i}{I_i} confluent for every state $s \in I_i$. Ideally, we
could leverage \algoref{InteractiveDecisionProcedure}, invoking it once per
segment. Unfortunately, \algoref{InteractiveDecisionProcedure} can only be used
to determine if $O$ is \sticonfluent{s}{T_i}{I_i} for a \emph{particular} state
$s \in I_i$, not for \emph{every} state $s \in I_i$. Thus, we would have to
invoke \algoref{InteractiveDecisionProcedure} $|I_i|$ times fore every segment
$(I_i, T_I)$, which is clearly infeasible given that $I_i$ can be large or even
infinite.

Instead, we develop a new decision procedure that can be used to determine if
an object is \sticonfluent{s}{T}{I} for every state $s \in I$. To do so, we
need to extend the notion of reachability to a notion of coreachability and
then generalize \thmref{IclosureEquivalentIconfluence}. Two states $s_1, s_2
\in I$ are \defword{coreachable} with respect to $T$ and $I$, abbreviated
\defword{$\TIcoreachable{s_1, s_2}$}, if there exists some state $s_0 \in I$
such that $s_1$ and $s_2$ are both $\sTIreachable{}$.

\begin{theorem}\thmlabel{ExtendedIclosureEquivalentIconfluence}
  Consider an object $O = (S, \join)$, a set of transactions $T$, and an
  invariant $I$. If every pair of states in the invariant are coreachable (i.e.
  $\forall s_1, s_2 \in S.\ \TIcoreachable{s_1, s_2}$), then
  \[
    \text{$O$ is \Iclosed{}}
    \iff
    \text{$O$ is \sticonfluent{s}{T}{I} for every $s \in I$}
  \]
\end{theorem}
\begin{proof}
  The proof of the forward direction is exactly the same as the proof of
  \thmref{IclosureImpliesIconfluence}. Transactions always maintain the
  invariant, so if merge does as well, then every reachable state must satisfy
  the invariant.
  %
  For the reverse direction, consider two arbitrary states $s_1, s_2 \in I$.
  The two points are $\TIcoreachable{}$, so there exists some state $s_0$ from
  which they can be reached. $O$ is \sTIconfluent{} and $s_1 \join s_2$ is
  $\sTIreachable{}$, so it satisfies the invariant.
\end{proof}

Using \thmref{ExtendedIclosureEquivalentIconfluence}, we develop
\algoref{ArbitraryStartInteractiveDecisionProcedure}: a natural generalization
of \algoref{InteractiveDecisionProcedure}.
\algoref{InteractiveDecisionProcedure} iteratively refines the set of
\emph{reachable} states whereas
\algoref{ArbitraryStartInteractiveDecisionProcedure} iteratively refines the
set of \emph{coreachable} states, but otherwise, the core of the two algorithms
is the same.\footnote{%
  Another small difference is that \IsIclosed{} behaves differently in
  \algoref{InteractiveDecisionProcedure} and
  \algoref{ArbitraryStartInteractiveDecisionProcedure}. In
  \algoref{ArbitraryStartInteractiveDecisionProcedure}, \IsIclosed{} returns a
  triple (closed, $s_1$, $s_2$) where if closed is false, then $s_1, s_2 \in I$
  are two states not in $NR$ such that $I(s_1)$ and $I(s_2)$ but $\lnot I(s_1
  \join s_2)$. If no such states exist, then closed is true, and $s_1$ and
  $s_2$ are null.
}
Now, a segmented invariant-confluence decision procedure, can simply invoke
\algoref{ArbitraryStartInteractiveDecisionProcedure} once on each segment.

\newcommand{\algocomment}[1]{\State \textcolor{flatdenim}{\texttt{//} #1}}
\begin{algorithm}[t]
  \caption{%
    Interactive invariant-confluence decision procedure for arbitrary start
    state $s \in I$
  }%
  \algolabel{ArbitraryStartInteractiveDecisionProcedure}
  \begin{algorithmic}
    \algocomment{Return if $O$ is \sticonfluent{s}{T}{I} for every $s \in I$.}
    \Function{IsInvConfluent}{$O$, $T$, $I$}
      \State \Return \Call{Helper}{$O$, $T$, $I$, $\emptyset$, $\emptyset$}
    \EndFunction

    \State

    \algocomment{$R$ is a set of $\TIcoreachable{}$ states.}
    \algocomment{$NR$ is a set of $\TIcounreachable{}$ states.}
    \Function{Helper}{$O$, $T$, $I$, $R$, $NR$}
      \State closed, $s_1$, $s_2$ $\gets$ \Call{IsIclosed}{$O$, $I$, $NR$}
      \If {closed}
        \State \Return true
      \EndIf
      \State Augment $R$ and $NR$ with random search and user input
      \If{$(s_1, s_2) \in R$}
        \Return false
      \Else\
        \Return \Call{Helper}{$O$, $T$, $I$, $R$, $NR$}
      \EndIf
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\begin{example}\examplelabel{CounreachableExample}
  Let $O = (\ints^3 \times \ints^3, \join)$ be a
  PN-Counter~\cite{shapiro2011comprehensive} replicated on $3$ machines. Every
  state $s = (p_1, p_2, p_3), (n_1, n_2, n_3)$ represents the integer $(p_1 +
  p_2 + p_3) - (n_1 + n_2 + n_3)$. To increment and decrement the counter, the
  $i$th server increments $p_i$ and $n_i$ respectively, and $\join$ computes an
  element-wise maximum. Our start state $s_0 = (0, 0, 0), (0, 0, 0)$; our set
  $T$ of transactions consists of increment and decrement; and our invariant
  $I$ is that the value of $s$ is non-negative.

  Applying \algoref{InteractiveDecisionProcedure}, \IsIclosed{} returns false
  with the states $s_1 = (1, 0, 0), (0, 1, 0)$ and $s_2 = (1, 0, 0), (0, 0,
  1)$. Both are reachable, so $O$ is not \sTIconfluent{} and
  \algoref{InteractiveDecisionProcedure} returns false.  The culprit is
  concurrent decrements, which we can forbid in a simple one-segment
  segmentation $\Sigma = (I, T^+)$ where $T^+$ consists only of increment
  transactions. Applying, \algoref{ArbitraryStartInteractiveDecisionProcedure},
  \IsIclosed{} again returns false with states $s_1$ and $s_2$. This time,
  however, the user recognizes that the two states are not
  $\ticoreachable{T^+}{I}{}$. The user refines $NR$ with the observation that
  two states are coreachable if and only if they have the same values of $n_1,
  n_2, n_3$. After this, \IsIclosed{} and
  \algoref{ArbitraryStartInteractiveDecisionProcedure} return true.
\end{example}

\subsection{Discussion}
There are a couple of things worth noting about segmented invariant-confluence,
its system model, and its decision procedure. First, invariant-confluence is a
very coarse-grained property. If an object is invariant-confluent, then we can
replicate it with no coordination. If it is not invariant-confluent, then we
have no guarantees. There's no in-between. Segmented invariant-confluence, on
the other hand, is a much more fine-grained property that can be applied to
applications with varying degrees of complexity. Segmented
invariant-confluence provides guarantees to complex applications that require a
large number of segments and to simple applications with a smaller number of
segments, whereas invariant-confluence only provides guarantees to applications
that can be segmented into a single segment.

Second, segmented invariant-confluence naturally subsumes a distributed locking
approach to replicating non-invariant-confluent objects.  This approach first
recognizes which transactions cannot be safely executed concurrently and then
requires them to acquire a distributed lock before
executing~\cite{balegas2015putting, gotsman2016cause}. For example, in a
banking application with the invariant that all balances remain non-negative,
concurrent deposits are permitted, but concurrent withdrawals can lead to
invariant violations. Thus, we require that withdrawals acquire a distributed
lock before executing. This example is exactly the same as
\exampleref{CounreachableExample} which we handled by simply removing
withdrawal transactions from our segmentation's set of transactions.

Third, we can integrate a couple of optimizations into our system model to
further reduce the amount of coordination required. First, if a server with
state $s_i$ and active segment $(I_i, T_i)$ receives a transaction $t \in I_i$
to execute, and $t(s_i)$ violates the active invariant but not the global
invariant, instead of initiating a round of global coordination, $p_i$ can
simply buffer $t$ for re-execution at a later time. While this increases the
latency required to execute $t$, it's possible that after other transactions
are executed, re-executing $t$ may lead to a state that either satisfies the
active invariant or violates the global invariant. In either case, a round of
global coordination is avoided. Similarly, servers can buffer transactions that
require global coordination, executing an entire batch of these transactions
during a single round of global coordination.

Fourth, a segmented invariant-confluence decision procedure can also leverage
\thmref{LatticeProperty} in addition to
\algoref{ArbitraryStartInteractiveDecisionProcedure}. If an object $O$ meets
criteria (1) - (3), then it is \sticonfluent{s}{T}{I} for every state $s \in
I$.

\todo{Forward reference examples in the evaluation section.}