### 2018-04-18 doc/paper/sections/introduction.tex
Work on related work.

\section{Introduction}
When an application designer decides to replicate a piece of data, they have to
make a fundamental choice between weak and strong consistency. Replicating the
data with strong consistency using a technique like state machine
replication~\cite{schneider1990implementing, lamport1998part,
liskov2012viewstamped, ongaro2014search, moraru2013there, vincent2015designing,
oki1988viewstamped, lamport2005generalized, lamport2006fast} makes the
application designer's life very easy. To the application developer, a strongly
consistent system behaves exactly like a single-threaded system running on a
single node, so reasoning about the behavior of the system is
simple~\cite{herlihy1990linearizability}.  Unfortunately, strong consistency is
at odds with performance. The CAP theorem and PACELC theorem tell us that
strongly consistent systems suffer from higher latency at best and
unavailability at worst~\cite{gilbert2002brewer, brewer2012cap,
abadi2012consistency}. On the other hand, weak consistency models like
eventual consistency~\cite{vogels2009eventually}, PRAM consistency~\cite{lipton1988pram},
causal consistency~\cite{ahamad1995causal}, and others~\cite{lloyd2011don,
mehdi2017can} allow data to be replicated with high availability and low
latency, but they put a tremendous burden on the application designer to reason
about the complex interleavings of operations that are allowed by these weak
consistency models. In fact, weak consistency models strip an application
developer of one of the earliest and most effective tools that is used to
reason about the execution of programs: application
invariants~\cite{hoare1969axiomatic, balegas2015towards}. Even if every
transaction executing in a weakly consistent system individually maintains an
application invariant, the system as a whole can produce invariant-violating
states.

Is it possible for us to have our strongly consistent cake and eat it with high
availability too? Can we replicate a piece of data with weak consistency but
still ensure that its invariants are maintained? Yes, sometimes. Bailis et al.\
introduced the notion of invariant-confluence as a necessary and sufficient
condition for when invariants can be maintained over replicated data without
the need for any coordination~\cite{bailis2014coordination}. If an object is
invariant-confluent with respect to an invariant, we can replicate it with the
performance benefits of weak consistency and (some of) the correctness benefits
of strong consistency. However, the task of identifying whether or not an
object actually is invariant-confluent remains a challenge. Bailis et al.\
manually categorized a set of common objects, transactions, and invariants
(e.g.\ foreign key constraints on relations, linear constraints on integers) as
invariant-confluent or not, but ideally we would have a general purpose program
that can automatically determine invariant-confluence for us. \textbf{The first
main thrust of this paper is to make progress towards a general-purpose
invariant-confluence decision procedure.}

Designing a general-purpose invariant-confluence decision procedure is
difficult\footnote{Actually, it's impossible. Determining whether an object is
invariant-confluent is, in general, undecidable. Still, we can develop a
decision procedure that works well in the common case.}. Many existing
approaches instead develop a decision procedure for invariant-closure, a
sufficient (but not necessary) condition for
invariant-confluence~\cite{li2012making, li2014automating}. The existing
approaches check to see if an object is invariant-closed. If it is, then they
conclude that it is also invariant-confluent. If it's not, then the current
approaches are unable to conclude anything about whether or not the object is
invariant-confluent. In this paper, we take a step back and study the
underlying reason \emph{why} invariant-closure is not necessary for
invariant-confluence. We find that invariant-confluence is fundamentally a
property about reachability, about whether certain values of our replicated
object can actually be obtained in the execution of a system. We discover that
unreachable, yet invariant-satisfying, states cause invariant-closure to be an
unnecessary condition. Using this understanding, we construct a set of modest
conditions under which invariant-closure and invariant-confluence are in fact
\emph{equivalent}, allowing us to reduce the problem of determining
invariant-confluence to that of determining invariant-closure.

Then, we use these conditions to design a general-purpose interactive
invariant-confluence decision procedure. Users iteratively interact with the
decision procedure, eliminating unreachable states from the invariant, moving
towards the conditions under which invariant-closure and invariant-confluence
are equivalent. Meanwhile, the decision-procedure generates example reachable
and unreachable states which help the user recognize patterns that describe
reachability. We also leverage our intuitions about reachability to develop a
new sufficient condition for invariant-confluence, dubbed merge-reducibility.
Merge-reducibility covers some cases that are not covered by invariant-closure,
and it can be automatically checked without any user interaction.

\textbf{The second main thrust of this paper is to take a step beyond
invariant-confluence and develop a generalization of invariant-confluence
called segmented invariant-confluence.} While invariant-confluence
characterizes objects which can be replicated \emph{without any} coordination,
segmented invariant-confluence allows us to replicate non-invariant-confluent
objects with \emph{minimal coordination}. The main idea is that
invariant-satisfying states are divided into a number of segments, and the set
of allowable of transactions within each segment is restricted. Within a
segment, servers act without any coordination, synchronizing only to transition
across segment boundaries. This design highlights the trade-off between
application complexity and coordination-freedom; more complex applications
require more segments which require more coordination, and vice-versa.

In closing, here is an outline of the paper and of our contributions:
\begin{itemize}
  \item
    We propose a novel expression-oriented definition of invariant-confluence
    that is both formal and simple (\secref{InvariantConfluence}).

  \item
    We develop an understanding of why invariant-closure is not necessary and
    use this understanding to develop conditions under which it is both
    necessary and sufficient (\secref{InvariantClosure}).

  \item
    We exploit these conditions to design a general-purpose interactive
    decision procedure for invariant-confluence
    (\secref{InteractiveDecisionProcedure}).

  \item
    We again exploit these conditions to design a novel non-trivial sufficient
    condition for invariant-confluence, merge-reducibility, that does not
    depend on user interaction (\secref{MergeReduction}).

  \item
    We present segmented invariant-confluence: a generalization of
    invariant-confluence that uses a small amount of coordination to maintain
    invariants for replicated objects that are otherwise not
    invariant-confluent (\secref{SegmentedInvariantConfluence}).

  \item
    We implement our decision procedures and sufficient conditions for both
    invariant-confluence and segmented invariant-confluence in a system called
    Lucy, then use Lucy to evaluate the efficacy or our methods
    (\secref{Evaluation}).
\end{itemize}