### 2018-04-22 doc/paper/sections/conclusion.tex
Tidied front matter + acks.

\section{Conclusion}
TODO

\textbf{Acknowledgments.}
The authors would like to thank Peter Alvaro, Alvin Cheung, Anthony Tan,
and Cristina Teodoropol for fruitful discussion and feedback.
%
This research is supported in part by DHS Award HSHQDC-16-3-00083, NSF CISE
Expeditions Award CCF-1139158, and gifts from Alibaba, Amazon Web Services, Ant
Financial, CapitalOne, Ericsson, GE, Google, Huawei, Intel, IBM, Microsoft,
Scotiabank, Splunk and VMware.