### 2018-04-22 doc/paper/sections/conclusion.tex
First stab at conclusion.

\section{Conclusion}
This paper revolved around two major contributions.  First, we developed a
deeper understanding of invariant-closure and invariant-confluence by looking
at the two criteria with reachability in mind. We found that invariant-closure
fails to incorporate a notion of reachability, and using this intuition, we
developed conditions under which invariant-closure and invariant-confluence are
equivalent. Second, we proposed a new consistency model and generalization of
invariant-confluence, segmented invariant-confluence, that can be used to
replicate non-invariant-confluent objects with a small amount of coordination
while still preserving their invariants. We also found that segmented
invariant-confluence naturally subsumes existing techniques for maintaining
invariants of replicated objects (e.g.\ locking and escrow transactions).

\textbf{Acknowledgments.}
The authors would like to thank Peter Alvaro, Alvin Cheung, Anthony Tan,
and Cristina Teodoropol for fruitful discussion and feedback.
%
This research is supported in part by DHS Award HSHQDC-16-3-00083, NSF CISE
Expeditions Award CCF-1139158, and gifts from Alibaba, Amazon Web Services, Ant
Financial, CapitalOne, Ericsson, GE, Google, Huawei, Intel, IBM, Microsoft,
Scotiabank, Splunk and VMware.