### 2018-04-22 doc/paper/sections/conclusion.tex

\section{Conclusion}
This paper revolved around two major contributions.  First, we developed a
deeper understanding of invariant-closure and \invariantconfluence{} by looking at the
two criteria with reachability in mind. We found that invariant-closure fails
to incorporate a notion of reachability, and using this intuition, we developed
conditions under which invariant-closure and \invariantconfluence{} are equivalent.
Second, we proposed a new consistency model and generalization of
\invariantconfluence{}, segmented \invariantconfluence{}, that can be used to replicate
non-invariant-confluent objects with a small amount of coordination while still
preserving their invariants. We also found that segmented \invariantconfluence{}
naturally subsumes existing techniques for maintaining invariants of replicated
objects (e.g.\ locking and escrow transactions).

\textbf{Acknowledgments.}
The authors would like to thank Peter Alvaro, Alvin Cheung, Anthony Tan,
and Cristina Teodoropol for fruitful discussion and feedback.
%
This research is supported in part by DHS Award HSHQDC-16-3-00083, NSF CISE
Expeditions Award CCF-1139158, and gifts from Alibaba, Amazon Web Services, Ant
Financial, CapitalOne, Ericsson, GE, Google, Huawei, Intel, IBM, Microsoft,
Scotiabank, Splunk and VMware.