### 2018-04-22 doc/paper/sections/merge_reduction.tex

\section{Merge Reduction}\seclabel{MergeReduction}
In \secref{InvariantClosure}, we discussed how \invariantclosure{} is
fundamentally a property of reachability and that \invariantclosure{} is
sufficient but not necessary for \invariantconfluence{} because it fails to
incorporate any notion of reachability. Using this intuition, we established
\thmref{IclosureEquivalentIconfluence} and then exploited the theorem in
\algoref{InteractiveDecisionProcedure}. In this section, we again take
advantage of this intuition to develop a new sufficient condition for
\invariantconfluence{} that can be checked without user interaction and that
covers some cases not covered by \invariantclosure{}.

An expression $e = t_1(t_2(\ldots(t_n(s))\ldots))$ is \defword{merge-free} if
does not contain any merges (i.e.\ it is generated by the grammar $e ::= s \mid
t(e)$). An object $O = (S, \join)$ is \defword{merge-reducible} with respect to
a start state $s_0 \in S$, a set of transactions $T$, and an invariant $I$,
abbreviated \defword{\sTImergereducible}, if for every pair of $e_1$ and $e_2$
of merge-free \sTIreachable{} expressions, there exists some merge-free
\sTIreachable{} expression $e_3$ that evaluates to the same state as $e_1
\join e_2$. Intuitively, if $O$ is merge-reducible, we can replace $e_1 \join
e_2$ (which has one merge) with $e_3$ (which has no merges) to obtain an
equivalent expression with fewer merges.

\begin{theorem}\thmlabel{ReducibilityImpliesIconfluence}
  Given an object $O = (S, \join)$, a start state $s_0 \in S$, a set of
  transactions $T$, and an invariant $I$, if $I(s_0)$ and if $O$ is
  \sTImergereducible{}, then $O$ is \sTIconfluent{}.
\end{theorem}

The proof of \thmref{ReducibilityImpliesIconfluence} is a straightforward
induction. We begin with an \sTIreachable{} expression $e$ and repeatedly
replace any subexpression that merges two merge-free subexpressions with an
equivalent merge-free reachable subexpression (which we can do because $O$ is
merge-reducible). We repeat this process until $e$ has been completely replaced
with an equivalent merge-free reachable expression $e'$. Because $I(s_0)$ and
because our system model only executes transactions that preserve the
invariant, $e'$ (and hence $e$) is guaranteed to satisfy the invariant. Thus,
all reachable states satisfy the invariant, so $O$ is \invariantconfluent{}.

Merge-reducibility is a sufficient condition for \invariantconfluence{}, but
unlike with \invariantclosure{}, it is not straightforward to automatically
determine if an object is merge-reducible. In \thmref{LatticeProperty}, we
outline a sufficient condition for merge-reducibility that is straightforward
to determine automatically.

\begin{theorem}\thmlabel{LatticeProperty}
  Given an object $O = (S, \join)$, a start state $s_0 \in S$, a set of
  transactions $T$, and an invariant $I$, if the following criteria are met,
  then $O$ is \sTIconfluent{}.
  \begin{enumerate}
    \item
      $O$ is a semilattice.

    \item
      For every $t \in T$, there exists some $s_t \in S$ such that for all $s
      \in S$, $t(s) = s \join s_t$. That is, every transaction $t$ is of the
      form $t(s) = s \join s_t$ for some constant $s_t$.

    \item
      For every pair of transactions $t_1, t_2 \in T$ and for all states $s \in
      S$, if $I(s)$, $I(t_1(s))$, and $I(t_2(s))$, then $I(t_1(s) \join
      t_2(s))$.

    \item
      $I(s_0)$.
  \end{enumerate}
\end{theorem}

The proof of \thmref{LatticeProperty} involves proving that if criteria (1)
through (4) are met, then $O$ is \sTImergereducible{}. From this, it's
immediate that $O$ is \sTIconfluent{} by
\thmref{ReducibilityImpliesIconfluence}. To show that $O$ is
\sTImergereducible, we consider two merge-free reachable expressions
\[
  e_1 = t_1(\ldots(t_n(s_0))\ldots)
  \quad\text{and}\quad
  e_2 = u_1(\ldots(u_m(s_0))\ldots)
\]
Then, we inductively use criterion (3) with criteria (1) and (2) to show that
\[
  e_3 = t_1(\ldots(
          t_n(u_1(\ldots(u_m(s_0))\ldots))
        )\ldots)
\]
is reachable and evaluates to $e_1 \join e_2$.

\thmref{IclosureImpliesIconfluence} states that \invariantclosure{} is a
sufficient condition for \invariantconfluence{}, and \thmref{LatticeProperty}
states that criteria (1) -- (4) are sufficient conditions for
\invariantconfluence{}. How do these sufficient conditions relate to one another?
Clearly, not all \invariantclosed{} objects are semilattices, so
\invariantclosure{} does not imply criteria (1) -- (4). Conversely, there are some
objects which satisfy criteria (1) -- (4) which are not \invariantclosed{}. Here's
one example.

\begin{example}\examplelabel{TwoSets}
  Let $O = (\mathcal{P}(\nats), \cup)$ where $\mathcal{P}(\nats)$ is the power
  set of the natural numbers. Our start state $s_0 = \set{0}$ is the set of
  $0$. Let $t_Y(X) = X \cup Y$ be the transaction which unions $Y$ with its
  argument $X$. Our set $T = \setst{t_Y}{Y \subseteq \nats}$ of transactions
  consists of all possible $t_Y$. Our invariant $I$ consists of all non-empty
  sets $X$ that contain only even or only odd elements. That is, $I = \setst{X
  \subseteq 2\nats}{X \neq \emptyset} \cup \setst{X \subseteq 2\nats + 1}{X
  \neq \emptyset}$.

  Clearly, criteria (1), (2), and (4) are satisfied. Criterion (3) is also
  satisfied. If $s$ is an arbitrary state that satisfies $I$, then it is
  non-empty and contains, without loss of generality, only even integers. If
  $t_1$ and $t_2$ are arbitrary transactions such that $I(t_1(s))$ and
  $I(t_2(s))$, then $t_1(s)$ and $t_1(s)$ are also non-empty and contain only
  even integers. Thus, $t_1(s) \cup t_2(s)$ is clearly non-empty and contains
  only even integers, so $I(t_1(s) \join t_2(s))$.
  %
  However, $O$ is not \Iclosed{}. Let $s_1 = \set{0}$ and $s_2 = \set{1}$.
  Then, $I(s_1)$ and $I(s_2)$, but letting $s_3 = s_1 \cup s_2 = \set{0, 1}$,
  $\lnot I(s_3)$.
\end{example}

{\input{figures/merge_reducible}}

\Invariantclosure{} is not necessary for \invariantconfluence{} because it fails to
incorporate any notion of reachability. Criteria (1) -- (4) are also
unnecessary, but they can be used to prove that some non-\invariantclosed{}
objects are \invariantconfluent{} because the criteria \emph{do} incorporate
notions of reachability. In particular, criterion (3) is a slight variant of
\invariantclosure{}; it also states that invariant satisfying states should be
closed under merge. The fundamental difference is that criterion (3) restricts
its attention to the merge of two states which are \emph{reachable} from a
common ancestor state.

In \exampleref{TwoSets}, we saw this fundamental difference rear its head. $O$
is not $I$-closed because the union of an odd-only set with an even-only set is
a set with both odd and even integers. However, if we begin in an
invariant satisfying state, we cannot reach both an odd-only and even-only set.
Criterion (3) is able to recognize this fact and conclude that $O$ is
\invariantconfluent{} despite it not being \invariantclosed{}.

% TODO: Discuss generality?