### 2018-04-22 doc/paper/sections/conclusion.tex
Finished paper tweaks.

\section{Conclusion}
This paper revolved around two major contributions.  First, we developed a
deeper understanding of \invariantclosure{} and \invariantconfluence{} by
looking at the two criteria with reachability in mind. We found that
\invariantclosure{} fails to incorporate a notion of reachability, and using
this intuition, we developed conditions under which \invariantclosure{} and
\invariantconfluence{} are equivalent.  Second, we proposed a new consistency
model and generalization of \invariantconfluence{}, segmented
\invariantconfluence{}, that can be used to replicate non-\invariantconfluent{}
objects with a small amount of coordination while still preserving their
invariants. We also found that segmented \invariantconfluence{} naturally
subsumes existing techniques for maintaining invariants of replicated objects
(e.g.\ locking and escrow transactions).

\textbf{Acknowledgments.}
The authors would like to thank Peter Alvaro, Alvin Cheung, Anthony Tan, and
Cristina Teodoropol for fruitful discussion and feedback.
%
This research is supported in part by DHS Award HSHQDC-16-3-00083, NSF CISE
Expeditions Award CCF-1139158, and gifts from Alibaba, Amazon Web Services, Ant
Financial, CapitalOne, Ericsson, GE, Google, Huawei, Intel, IBM, Microsoft,
Scotiabank, Splunk and VMware.