### 2018-05-01 doc/paper/sections/abstract.tex
took a pass

\begin{abstract}
Strongly consistent systems are easy to reason about but face fundamental
limitations in availability and performance. Weakly consistent systems can be
implemented with very high performance but place a burden on the application
developer to reason about complex interleavings of execution.
\Invariantconfluence{} provides a formal framework for understanding when we
can get the best of both worlds. An \invariantconfluent{} object can be
efficiently replicated with no coordination needed to preserve its invariants.
However, actually determining whether or not an object is \invariantconfluent{}
is challenging: undecidable in general. In this paper, we establish conditions under
which a commonly used sufficient condition for \invariantconfluence{} is both
necessary and sufficient and use this condition to design (a) a general-purpose
interactive \invariantconfluence{} decision procedure and (b) a novel
sufficient condition that can be checked automatically. We then take a step
beyond \invariantconfluence{} and introduce a generalization of
\invariantconfluence{} called segmented \invariantconfluence{}, which allows us
to replicate non-\invariantconfluent{} objects with a small amount of
coordination. We implemented our theoretical findings in a prototype called
Lucy and found that our decision procedures efficiently handle common real-world workloads
including foreign key constraints, rollups, escrow transactions, and others.  We also
found that segmented \invariantconfluent{} replication can deliver up to an
order of magnitude more throughput than linearizable replication for low
contention workloads and comparable throughput for medium to high contention
workloads.
\end{abstract}