### 2018-09-19 doc/paper/sections/related_work.tex
Added CRDT section to related work.
From reviewer 11:

> A clear connection to what CRDTs are within scope would improve the
> contribution of this paper.

and

> You make a connection to Shapiro et al.'s work on CRDT and use a few
> examples that are popular (e.g. PN-Counter), but can you go further
> and connect to more complex structures like Sequence. This will allow
> you to reach a broader audience and let them clearly understand the
> broader implication of this work. (BTW, nice discussion of PK-FK).

Previously, we had a paragraph on CRDTs in the related works section. I
cut it out bec

\section{Related Work}
% Other possibilities:
%   - Towards Fast Invariant Preservation in Geo-replicated Systems
%   - Feral Concurrency Control
%   - The CISE Tool: Proving Weakly-Consistent Applications Correct
%   - Declarative Programming over Eventually Consistent Data Stores
%   - Extending Eventually Consistent Cloud Databases for Enforcing Numeric Invariants
% In this section, we detail four alternate approaches to maintaining invariants
% without coordination: RedBlue consistency and Sieve, the homeostasis and
% demarcation protocols, explicit consistency, and token based invariant
% confluence. We then survey other bodies of related work.

\newcommand{\sieve}{\textsc{SIEVE}}
\textbf{RedBlue Consistency and \sieve.}
RedBlue consistency is a consistency model that sits between causal consistency
and linearizability~\cite{li2012making}.
%
With RedBlue consistency, every operation is manually labelled as either red or
blue. All operations are executed with causal consistency, but with the added
restrictions that red operations are executed in a single total order embedded
within the causal ordering.
%
In~\cite{li2012making}, Li et al.\ introduce invariant safety as a
sufficient (but not necessary) condition for RedBlue consistent objects to be
\invariantconfluent{}. Invariant safety is an analog of \invariantclosure{}.
In~\cite{li2014automating}, Li et al.\ develop sophisticated techniques for
deciding invariant safety that involve calculating weakest preconditions. These
techniques are complementary to our work and can be used to improve the
\invariantclosure{} subroutine used by our decision procedures.
%
In contrast with these techniques, our \invariantconfluence{} decision
procedures can determine the \invariantconfluence{} of objects that are
\emph{not} invariant safe.

\textbf{The Demarcation and Homeostasis Protocols.}
The homeostasis protocol~\cite{roy2015homeostasis}, a generalization of the
demarcation protocol~\cite{barbara1994demarcation}, uses program analysis to
avoid unnecessary coordination between servers in a \emph{sharded} database
(whereas \invariantconfluence{} targets \emph{replicated} databases).  The
protocol guarantees that transactions are executed with observational
equivalence with respect to some serial execution of the transactions. This
means that intermediate states may be inconsistent, but externally observable
side effects and the final database state are consistent.  The observational
equivalence guaranteed by the homeostasis protocol is stronger than the
guarantees of \invariantconfluence{}. As a result, there are invariants and
workloads that the homeostasis protocol would execute with more coordination
than a segmented \invariantconfluent{} execution. Moreover, the homeostasis and
demarcation protocols' mechanism of establishing global invariants and
operating without coordination so long as the invariants are maintained is very
similar to our design of segmented \invariantconfluence{}.

\textbf{Explicit Consistency.}
Explicit consistency~\cite{balegas2015towards} is a consistency model that
combines \invariantconfluence{} and causal consistency, similar to RedBlue
consistency with invariant safety. To determine if a workload is amenable to
explicitly consistent replication, Balegas et al.\ determine if all pairs of
transactions can be concurrently executed on the same start state without
violating the invariant~\cite{balegas2015towards}. Balegas et al.\ argue that
this is a sufficient condition for explicit consistency. It is similar to
criterion (3) in \thmref{LatticeProperty}. In our work, we take a step further
and explore sufficient \emph{and necessary} conditions for
\invariantconfluence{}.
%
Balegas et al.\ also describe a variety of techniques---like conflict
resolution, locking, and escrow transactions~\cite{o1986escrow}---that can be
used to replicate workloads that do not meet their sufficient conditions.
Segmented \invariantconfluence{} is a general-purpose formalism that can be
used to model simple forms of these techniques.

\textbf{Token Based Invariant Confluence.}
In~\cite{gotsman2016cause}, Gotsman et al.\ discuss a hybrid token based
consistency model that generalizes a family of consistency models including
causal consistency, sequential consistency, and RedBlue consistency. An
application designer defines a set of tokens and specifies which pairs of
tokens conflict, and transactions acquire some subset of the tokens when they
execute. This allows the application designer to specify which transactions
conflict with one another.  Gotsman et al.\ develop sufficient conditions to
determine whether a given token scheme is sufficient to guarantee that a global
invariant is never broken. This model of consistency is very similar to
segmented \invariantconfluence{}. The token based approach allows users to
specify certain conflicts that are not possible with segmented
\invariantconfluence{}. A segmentation only allows transactions within a
segment to acquire a single self-conflicting lock. However, segmented
\invariantconfluence{} also introduces the notion of invariant segmentation,
which cannot be emulated with the token based approach. For example, it is
difficult to emulate escrow transactions with the token based approach.

\textbf{Serializable Distributed Databases.}
In \secref{Evaluation}, we saw that segmented \invariantconfluent{} replication
vastly outperforms linearizable replication for low coordination workloads, and
it performs comparably or worse for medium and high coordination workloads.
Distributed databases like Calvin~\cite{thomson2012calvin},
Janus~\cite{mu2016consolidating}, and TAPIR~\cite{zhang2015building} employ
algorithmic optimizations to implement serializable transactions with high
throughput and low latency.
%
While segmented \invariantconfluent{} replication will likely always outperform
serializable replication for low coordination workloads, these databases make
serializable replication the most performant option for executing workloads
that require a modest amount of coordination.

\textbf{Branch and Merge.}
Bayou~\cite{terry1995managing}, Dynamo~\cite{decandia2007dynamo}, and
TARDiS~\cite{crooks2016tardis} all take a branch and merge approach to
maintaining distributed invariants without coordination. With this approach,
servers execute transactions without any coordination but keep track of the
causal dependencies between transactions. Periodically, two servers merge
states and invoke a user defined merge function to reconcile the divergent
states. This approach does \emph{not} provide any formal guarantees that
invariants are maintained. Its correctness depends on the correctness of the
potentially complex user defined merge functions.

\textbf{CRDTs.}
CRDTs~\cite{shapiro2011conflict, shapiro2011comprehensive} are distributed
semilattices with inflationary update methods. Due to their algebraic
properties, CRDTs can be replicated with strong eventual consistency without
the need for any coordination. Our definition of distributed objects and our
\invariantconfluence{} system model are inspired directly by the corresponding
definitions and system models in~\cite{shapiro2011conflict}.
%
CRDTs are eventually consistent but may not preserve invariants. Conversely,
\invariantconfluent{} objects preserve invariants but may not be eventually
consistent. Thus, it is natural (though not necessary) to use CRDTs as
distributed objects. If a CRDT is determined to be \invariantconfluent{} with
respect to a particular invariant and set of transactions, then it achieves a
combination of strong eventual consistency and invariant preservation. Any CRDT
(e.g., counters, sets, graphs, sequences) can be used for this purpose.
%
Finally, our criteria in \thmref{LatticeProperty} also borrow ideas from CRDTs,
exploiting the nice algebraic properties of semilattices.

\textbf{CALM Theorem.}
Bloom~\cite{alvaro2010boom, alvaro2011consistency, conway2012logic} and its
formalism, Dedalus~\cite{alvaro2011dedalus, alvaro2013declarative}, are
declarative Datalog-based programming languages that are designed to program
distributed systems. The accompanying CALM
theorem~\cite{hellerstein2010declarative, ameloot2013relational} states that if
and only if a program can be written in the monotone fragment of these
languages, then there exists a consistent, coordination-free implementation of
the program.  The CALM theorem provides guarantees about the consistency of
program outputs. It does not directly capture our notions of transactions or
invariant maintenance during program execution.  Moreover, Bloom and Dedalus
are general purpose programming languages that can be used to implement a
variety of distributed systems that are outside of the scope of
\invariantconfluence{}.

\textbf{Program Analysis in Database Systems.}
Using program analysis to improve the performance of database systems is not
new. For example, it has been used to improve the performance of
database-backed web applications~\cite{cheung2014using, wu2016transaction,
ramachandra2012program} and used to improve the performance of optimistic
concurrency control on multi-core machines~\cite{wu2016transaction}. Our work
on \invariantconfluence{} continues the theme of using program analysis to reap
the performance benefits gained from understanding program semantics.