### 2018-09-20 doc/paper/sections/abstract.tex
Discussed limitations + shrunk paper.
From reviewer 11:

> Discussion of limitations and practical use can be made clearer.

From reviewer 12:

> While the ideas of the paper are very nice, the paper may more
> concretely describe how practical the algorithms are. For example, the
> first decision procedure (Algorithm 1) involves human to augment NR.
> In general, what are the tricky classes of transactions that may
> require lot of human interactions? Similar comments apply to segmented
> invariant confluence (Algorithm 5). What classes of
> transactions/applications that may require complex

\begin{abstract}
Strongly consistent distributed systems are easy to reason about but face
fundamental limitations in availability and performance. Weakly consistent
systems can be implemented with very high performance but place a burden on the
application developer to reason about complex interleavings of execution.
\Invariantconfluence{} provides a formal framework for understanding when we
can get the best of both worlds. An \invariantconfluent{} object can be
efficiently replicated with no coordination needed to preserve its invariants.
However, actually determining whether or not an object is \invariantconfluent{}
is challenging: undecidable in general.

In this paper, we establish conditions under which a commonly used sufficient
condition for \invariantconfluence{} is both necessary and sufficient and use
this condition to design (a) a general-purpose interactive
\invariantconfluence{} decision procedure and (b) a novel sufficient condition
that can be checked automatically. We then take a step beyond
\invariantconfluence{} and introduce a generalization of \invariantconfluence{}
called segmented \invariantconfluence{}, which allows us to replicate
non-\invariantconfluent{} objects with a small amount of coordination.

We implemented these formalisms in a prototype called Lucy and found that our
decision procedures efficiently handle common real-world workloads including
foreign keys, rollups, escrow transactions, and more.  We also found that
segmented \invariantconfluent{} replication can deliver up to an order of
magnitude more throughput than linearizable replication for low contention
workloads and comparable throughput for medium to high contention workloads.
\end{abstract}