### 2018-09-20 doc/paper/sections/abstract.tex
Minor typo fixes, grammar fixes, and rewording.

\begin{abstract}
Strongly consistent distributed systems are easy to reason about but face
fundamental limitations in availability and performance. Weakly consistent
systems can be implemented with very high performance but place a burden on the
application developer to reason about complex interleavings of execution.
\Invariantconfluence{} provides a formal framework for understanding when we
can get the best of both worlds. An \invariantconfluent{} object can be
efficiently replicated with no coordination needed to preserve its invariants.
However, actually determining whether or not an object is \invariantconfluent{}
is challenging.

In this paper, we establish conditions under which a commonly used sufficient
condition for \invariantconfluence{} is both necessary and sufficient, and we use
this condition to design (a) a general-purpose interactive
\invariantconfluence{} decision procedure and (b) a novel sufficient condition
that can be checked automatically. We then take a step beyond
\invariantconfluence{} and introduce a generalization of \invariantconfluence{},
called segmented \invariantconfluence{}, that allows us to replicate
non-\invariantconfluent{} objects with a small amount of coordination.

We implemented these formalisms in a prototype called Lucy and found that our
decision procedures efficiently handle common real-world workloads including
foreign keys, rollups, escrow transactions, and more.  We also found that
segmented \invariantconfluent{} replication can deliver up to an order of
magnitude more throughput than linearizable replication for low contention
workloads and comparable throughput for medium to high contention workloads.
\end{abstract}