### 2019-08-15 doc/vldb_slides/iconfluence.tex
Added motivating bank example.

\documentclass[xcolor={dvipsnames,svgnames,table}]{beamer}

\usepackage{amsmath}
\usepackage{centernot}
\usepackage{etoolbox}
\usepackage{mathtools}
\usepackage{python}
\usepackage{slides}
\usepackage{stmaryrd}
\usepackage{tikz}
\usetikzlibrary{arrows}
\usetikzlibrary{backgrounds}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}

\title{Interactive Checks for Coordination Avoidance}
\author{Michael Whittaker, Joseph M. Hellerstein}
\date{}

\begin{document}
\begin{frame}
  \maketitle
\end{frame}

{\input{sections/four_wise_men.tex}}
{\input{sections/bank_example.tex}}
\end{document}