### 2019-08-16 doc/vldb_slides/sections/iclosure.tex
Added invariant closure notes.

\begin{frame}
  \Huge
  \begin{center}
    Goal: develop an invariant-confluence decision procedure.
  \end{center}

  \note{%
    Now that we've defined invariant confluence, we can turn our attention to
    the main goal of our paper. And that is to develop an invariant-confluence
    decision procedure. Determining whether or not a distributed object is
    invariant confluent is not easy to do by hand, so we'd like to develop a
    decision procedure that can automatically check whether something is
    invariant confluent for us.
  }
\end{frame}

\begin{frame}
  \Huge
  \begin{center}
    Reasoning about reachable states is hard.
  \end{center}

  \note{%
    Unfortunately, developing an invariant confluence decision procedure
    straight up is not easy. Invariant confluence is fundamentally a property
    about reachable states but reasoning automatically about reachable states
    is hard. I'm not saying it's impossible, but certainly I don't know how to
    do it.

    Instead of reasoning about invariant confluence directly then, let's look
    at a sufficient condition for invariant confluence called invariant
    closure.
  }
\end{frame}

\begin{frame}
  \Large
  We say a set $S$ is \defword{closed under $f$} if for every $x, y \in S, f(x,
  y) \in S$. \pause For example,
  \begin{itemize}
    \item Even numbers are closed under addition.
    \item Odd numbers are \emph{not} closed under addition.
  \end{itemize}

  \note{%
    First, a quick refresher on terminology. We say a set $S$ is closed under a
    binary operator $f$ if for every $x$ and $y$ in $S$, $f(x, y)$ is also in
    $S$. For example, even numbers are closed under addition because the sum of
    any two even numbers is even. But odd numbers are not closed under addition
    because the sum of two odd numbers is even.
  }
\end{frame}

\begin{frame}
  \Large
  $O = (S, \join)$ is \defword{\invariantclosed{}} with respect to an invariant
  $I$, abbreviated \defword{\Iclosed{}}, if invariant satisfying states are
  closed under merge. \pause That is, for every state $s_1, s_2 \in S$, if
  $I(s_1)$ and $I(s_2)$, then $I(s_1 \join s_2)$.

  \note{%
    We say that an object $O$ is invariant closed if invariant satisfying
    states are closed under merge. That is, if an object is invariant closed,
    then for every pair of states $s_1$ and $s_2$, if $s_1$ and $s_2$ satisfy
    the invariant then so does $s_1 \join s_2$.
  }
\end{frame}

\begin{frame}
  \Large
  \[
    \text{invariant closure} \implies \text{invariant confluence}
  \]
  \pause
  Transactions perserve invariants. If merging does too, then all reachable
  states satisfy the invariant.

  \note{%
    Invariant closure is a sufficient condition for invariant confluence. Why
    is that? Well, transactions already preserve invariants. So if merging does
    as well, then every reachable state always maintains the invariant.
  }
\end{frame}

\begin{frame}
  \Large
  Checking invariant closure is more straightforward.
  \pause
  \begin{align*}
    \forall x_1, & y_1, x_2, y_2.\, \\
    \quad & x_1y_1 \leq 0 \land x_2y_2 \leq 0 \implies \\
    \quad & \max(x_1, x_2)\max(y_1, y_2) \leq 0
  \end{align*}

  \note{%
    The good thing about invariant closure is that it's much easier to check
    automatically. for example, remember that example we had with pairs of
    integers? We can pose whether or not that object is invariant closed as
    this formula, and we can pass this formula directly to an SMT solver to
    figure out whether it's invariant closed.
  }
\end{frame}

\begin{frame}
  \Large
  \[
    \text{invariant closure} \implies \text{invariant confluence}
  \]

  \note{%
    So to decide whether an object is invariant confluent, we can take the
    object and ask if it's invariant closed. If it is, then it's also invariant
    confluent and we're done. If it's not invariant closed, then what do we
    know?
  }
\end{frame}

\begin{frame}
  \Large
  \[
    \text{invariant closure}
    \xLeftarrow{\phantom{a}?\phantom{a}}
    \text{invariant confluence}
  \]

  \note{%
    Well, that depends on whether invariant confluence implies invariant
    closure.
  }
\end{frame}

\newcommand{\xmin}{-2}
\newcommand{\xmax}{2}
\newcommand{\ymin}{-2}
\newcommand{\ymax}{2}

% Axes.
\newcommand{\xyaxes}{
  \draw[] (\xmin.5, 0) to (\xmax.5, 0);
  \draw[] (0, \ymin.5) to (0, \ymax.5);
  \node at (\xmax + 1, 0) {$x$};
  \node at (0, \ymax + 1) {$y$};
}

% Quadrant 1.
\newcommand{\quadi}[5]{{
  \newcommand{\argstyle}{#1}
  \newcommand{\argxmin}{#2}
  \newcommand{\argxmax}{#3}
  \newcommand{\argymin}{#4}
  \newcommand{\argymax}{#5}
  \foreach \x in {0, ..., \argxmax} {
    \foreach \y in {0, ..., \argymax} {
      \node[\argstyle] (\x-\y) at (\x, \y) {};
    }
  }
}}

% Quadrant 2.
\newcommand{\quadii}[5]{{
  \newcommand{\argstyle}{#1}
  \newcommand{\argxmin}{#2}
  \newcommand{\argxmax}{#3}
  \newcommand{\argymin}{#4}
  \newcommand{\argymax}{#5}
  \foreach \x in {\argxmin, ..., 0} {
    \foreach \y in {0, ..., \argymax} {
      \node[\argstyle] (\x-\y) at (\x, \y) {};
    }
  }
}}

% Quadrant 3.
\newcommand{\quadiii}[5]{{
  \newcommand{\argstyle}{#1}
  \newcommand{\argxmin}{#2}
  \newcommand{\argxmax}{#3}
  \newcommand{\argymin}{#4}
  \newcommand{\argymax}{#5}
  \foreach \x in {\argxmin, ..., 0} {
    \foreach \y in {\argymin, ..., 0} {
      \node[\argstyle] (\x-\y) at (\x, \y) {};
    }
  }
}}

% Quadrant 4.
\newcommand{\quadiv}[5]{{
  \newcommand{\argstyle}{#1}
  \newcommand{\argxmin}{#2}
  \newcommand{\argxmax}{#3}
  \newcommand{\argymin}{#4}
  \newcommand{\argymax}{#5}
  \foreach \x in {0, ..., \argxmax} {
    \foreach \y in {\argymin, ..., 0} {
      \node[\argstyle] (\x-\y) at (\x, \y) {};
    }
  }
}}

% State labels.
\newcommand{\statelabels}{
  \node[statelabel] at (0, 0) {$s_0$};
  \node[statelabel] at (-1, 1) {$s_1$};
  \node[statelabel] at (1, -1) {$s_2$};
  \node[statelabel] at (1, 1) {$s_3$};
}

\tikzstyle{point}=[shape=circle, fill=flatgray, inner sep=3pt]
\tikzstyle{inv}=[line width=0.75pt, draw=black]
\tikzstyle{pointinv}=[point, inv]
\tikzstyle{invregion}=[rounded corners, fill=flatgreen!50, draw=none]
\tikzstyle{reachableregion}=[rounded corners, fill=flatblue!50, draw=none]
\tikzstyle{statelabel}=[anchor=south west, inner sep=1pt]

\begin{frame}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \centering
      \begin{tikzpicture}[scale=1]
        \begin{scope}
          \clip (\xmin.5, \ymax.5) rectangle (\xmax.5, \ymin.5);
          \draw[invregion] (\xmin.9, \ymax.9) rectangle (0.5, -0.5);
          \draw[invregion] (-0.5, 0.5) rectangle (\xmax.9, \ymin.9);
          \draw (0.5, 0.5) to (0.5, \ymax.5);
          \draw (0.5, 0.5) to (\xmax.5, 0.5);
          \draw (-0.5, -0.5) to (-0.5, \ymin.5);
          \draw (-0.5, -0.5) to (\xmin.5, -0.5);
        \end{scope}

        \xyaxes{}
        \quadi{point}{\xmin}{\xmax}{\ymin}{\ymax}
        \quadiii{point}{\xmin}{\xmax}{\ymin}{\ymax}
        \quadii{pointinv}{\xmin}{\xmax}{\ymin}{\ymax}
        \quadiv{pointinv}{\xmin}{\xmax}{\ymin}{\ymax}
      \end{tikzpicture}

      {\Huge Invariant}
    \end{column}
    \begin{column}{0.5\textwidth}
      \centering
      \begin{tikzpicture}[scale=1]
        \begin{scope}
          \clip (-1, 1) rectangle (\xmax.5, \ymin.5);
          \draw[reachableregion, draw=black] (-0.5, 0.5) rectangle (\xmax.9, \ymin.9);
        \end{scope}

        \xyaxes{}
        \quadi{point}{\xmin}{\xmax}{\ymin}{\ymax}
        \quadiii{point}{\xmin}{\xmax}{\ymin}{\ymax}
        \quadii{point}{\xmin}{\xmax}{\ymin}{\ymax}
        \quadiv{pointinv}{\xmin}{\xmax}{\ymin}{\ymax}
      \end{tikzpicture}

      {\Huge Reachable}
    \end{column}
  \end{columns}

  \note{%
    To see if it does, let's revisit the example from before. Recall that in
    this example, our object is invariant confluent.
  }
\end{frame}

\begin{frame}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \centering
      \begin{tikzpicture}[scale=1]
        \begin{scope}
          \clip (\xmin.5, \ymax.5) rectangle (\xmax.5, \ymin.5);
          \draw[invregion] (\xmin.9, \ymax.9) rectangle (0.5, -0.5);
          \draw[invregion] (-0.5, 0.5) rectangle (\xmax.9, \ymin.9);
          \draw (0.5, 0.5) to (0.5, \ymax.5);
          \draw (0.5, 0.5) to (\xmax.5, 0.5);
          \draw (-0.5, -0.5) to (-0.5, \ymin.5);
          \draw (-0.5, -0.5) to (\xmin.5, -0.5);
        \end{scope}

        \xyaxes{}
        \quadi{point}{\xmin}{\xmax}{\ymin}{\ymax}
        \quadiii{point}{\xmin}{\xmax}{\ymin}{\ymax}
        \quadii{pointinv}{\xmin}{\xmax}{\ymin}{\ymax}
        \quadiv{pointinv}{\xmin}{\xmax}{\ymin}{\ymax}

        \draw[-latex, ultra thick] (-1, 2) to (2, 2);
        \draw[-latex, ultra thick] (2, -1) to (2, 2);
      \end{tikzpicture}

      {\Huge Invariant}
    \end{column}
    \begin{column}{0.5\textwidth}
      \centering
      \begin{tikzpicture}[scale=1]
        \begin{scope}
          \clip (-1, 1) rectangle (\xmax.5, \ymin.5);
          \draw[reachableregion, draw=black] (-0.5, 0.5) rectangle (\xmax.9, \ymin.9);
        \end{scope}

        \xyaxes{}
        \quadi{point}{\xmin}{\xmax}{\ymin}{\ymax}
        \quadiii{point}{\xmin}{\xmax}{\ymin}{\ymax}
        \quadii{point}{\xmin}{\xmax}{\ymin}{\ymax}
        \quadiv{pointinv}{\xmin}{\xmax}{\ymin}{\ymax}
      \end{tikzpicture}

      {\Huge Reachable}
    \end{column}
  \end{columns}

  \note{%
    But, note that our object is not invariant closed. The points $(-1, 2)$ and
    $(2, -1)$ both satisfy the invariant, but their merge does not.
  }
\end{frame}

\begin{frame}
  \Large
  \[
    \text{invariant closure} \centernot\impliedby \text{invariant confluence}
  \]

  \note{%
    Thus, our object is invariant confluent but not invariant closed, so
    invariant confluence does not imply invariant closure. This is unfortunate.
    We can check for invariant closure but not invariant confluence, so we'd
    like the two to be equivalent.

    Are there any situations in which the two do happen to be equivalent?
  }
\end{frame}

\begin{frame}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \centering
      \begin{tikzpicture}[scale=1]
        \begin{scope}
          \clip (\xmin.5, \ymax.5) rectangle (\xmax.5, \ymin.5);
          \draw[invregion] (\xmin.9, \ymax.9) rectangle (0.5, -0.5);
          \draw[invregion] (-0.5, 0.5) rectangle (\xmax.9, \ymin.9);
          \draw (0.5, 0.5) to (0.5, \ymax.5);
          \draw (0.5, 0.5) to (\xmax.5, 0.5);
          \draw (-0.5, -0.5) to (-0.5, \ymin.5);
          \draw (-0.5, -0.5) to (\xmin.5, -0.5);
        \end{scope}

        \xyaxes{}
        \quadi{point}{\xmin}{\xmax}{\ymin}{\ymax}
        \quadiii{point}{\xmin}{\xmax}{\ymin}{\ymax}
        \quadii{pointinv}{\xmin}{\xmax}{\ymin}{\ymax}
        \quadiv{pointinv}{\xmin}{\xmax}{\ymin}{\ymax}

        \draw[-latex, ultra thick] (-1, 2) to (2, 2);
        \draw[-latex, ultra thick] (2, -1) to (2, 2);
      \end{tikzpicture}

      {\Huge Invariant}
    \end{column}
    \begin{column}{0.5\textwidth}
      \centering
      \begin{tikzpicture}[scale=1]
        \begin{scope}
          \clip (-1, 1) rectangle (\xmax.5, \ymin.5);
          \draw[reachableregion, draw=black] (-0.5, 0.5) rectangle (\xmax.9, \ymin.9);
        \end{scope}

        \xyaxes{}
        \quadi{point}{\xmin}{\xmax}{\ymin}{\ymax}
        \quadiii{point}{\xmin}{\xmax}{\ymin}{\ymax}
        \quadii{point}{\xmin}{\xmax}{\ymin}{\ymax}
        \quadiv{pointinv}{\xmin}{\xmax}{\ymin}{\ymax}
      \end{tikzpicture}

      {\Huge Reachable}
    \end{column}
  \end{columns}

  \note{%
    Well, let's take a look at our example again. We noticed that our object is
    not invariant closed because we can merge $(-1, 2)$ and $(2, -1)$ to
    violate the invariant. But, notice that the point $(-1, 2)$ is not
    reachable.
  }
\end{frame}

\begin{frame}
  \Large
  If $I$ is a subset of reachable states, then
  \[
    \text{invariant closure} \iff \text{invariant confluence}
  \]
  \pause
  \begin{itemize}
    \item
      Forward direction: same as before.
    \item
      Backward direction: $I$ is equal to the set of reachable states.
      Reachable states are closed under merge, so so is $I$.
  \end{itemize}

  \note{%
    This example is actually representative of the overall trend. An invariant
    confluent object may not be invariant closed but only because we're merging
    points that are unreachable. If the invariant is a subset of the reachable
    states, then invariant closure and invariant confluence are identical.

    Why? Well, the forward direction here is what we proved before. For the
    backwards direction, we have that the invariant and the set of reachable
    points are identical. The set of reachable states is closed under merge, so
    so is the invariant, so the object is invariant closed.

    I know I'm walking through a lot of theory pretty quickly, but if you
    remember one slide, remember this one. If the invariant is a subset of the
    reachable states, then invariant closure and invariant confluence are
    identical.
  }
\end{frame}