### 2019-08-16 doc/vldb_slides/iconfluence.tex
Finished first cut at speaker notes.

\documentclass[xcolor={dvipsnames,svgnames,table}]{beamer}

\usepackage[noend]{algpseudocode}
\usepackage{algorithm}
\usepackage{amsmath}
\usepackage{centernot}
\usepackage{etoolbox}
\usepackage{invariantconfluence}
\usepackage{mathpartir}
\usepackage{mathtools}
\usepackage{pervasives}
\usepackage{pgfpages}
\usepackage{python}
\usepackage{slides}
\usepackage{stmaryrd}
\usepackage{tikz}
\usetikzlibrary{arrows}
\usetikzlibrary{backgrounds}
\usetikzlibrary{calc}
\usetikzlibrary{decorations.pathreplacing}
\usetikzlibrary{positioning}

\title{Interactive Checks for Coordination Avoidance}
\author{Michael Whittaker, Joseph M. Hellerstein}
\date{}

% https://gist.github.com/andrejbauer/ac361549ac2186be0cdb
% \setbeameroption{hide notes} % Only slides
% \setbeameroption{show only notes} % Only notes
\setbeameroption{show notes on second screen=right} % Both
\setbeamertemplate{note page}{\pagecolor{yellow!10}\insertnote}

\begin{document}
\begin{frame}
  \maketitle
\end{frame}

{\input{sections/four_wise_men.tex}}
{\input{sections/bank_example.tex}}
{\input{sections/iconfluence.tex}}
{\input{sections/iclosure.tex}}
{\input{sections/decision_procedure.tex}}
{\input{sections/segmented_iconfluence.tex}}
{\input{sections/evaluation.tex}}
\end{document}