### 2019-08-16 doc/vldb_slides/sections/evaluation.tex

\begin{frame}[fragile]
  \begin{verbatim}
checker = InteractiveInvariantConfluenceChecker()
x = checker.int_max('x', 0) # An int, x, merged by max.
y = checker.int_max('y', 0) # An int, y, merged by max.
checker.add_transaction('inc_x', [x.assign(x + 1)])
checker.add_transaction('dec_y', [y.assign(y - 1)])
checker.add_invariant(x * y <= 0)
checker.check()\end{verbatim}

  \note{%
    Finally, I'll give a teaser of our evaluation. We implemented our decision
    procedure in Python. You can specify objects, transactions, and invariants
    with code that looks like this.
  }
\end{frame}

\begin{frame}
  \begin{itemize}
    \item Foreign keys
    \item Auction application
    \item Escrow transactions
    \item TPC-C
  \end{itemize}

  \pause
  All run in less than a second and are implemented in less than 75 lines of
  specification.

  \note{%
    We specified a bunch of examples including things involving foreign keys,
    an auction application, something involving escrow transactions, and TPC-C.
    All the examples verify with a runtime of less then a second and every one
    is implemented in fewer than 75 lines of specification.
  }
\end{frame}

\begin{frame}
  \includegraphics[width=\textwidth]{assets/throughput_vs_fraction_16.pdf}

  \note{%
    We have some graphs showing that no coordination vastly outperforms
    replicating with strong consistency, and how segmented invariant confluence
    allows a smooth tradeoff between the two.
  }
\end{frame}

\begin{frame}
  \begin{center}
    \Huge
    Come to our poster tomorrow night!
  \end{center}

  \note{%
    I've jumped over a ton of details in the paper, and some things I didn't
    even mention at all. If you're curious about anything, I encourage you to
    come to our poster tomorrow night where I'm happy to chat.
  }
\end{frame}

\begin{frame}
  \begin{center}
    \Huge
    Thank you
  \end{center}

  \note{%
    Thank you all for listening, and I'm happy to take questions.
  }
\end{frame}