### 2019-08-16 doc/vldb_slides/sections/four_wise_men.tex

\tikzstyle{pic}=[inner sep=0pt]
\tikzstyle{quote}=[fill=gray!30, inner sep=12pt]

\newcommand{\vogels}{
  \node[pic] (vogels) at (0, 0) {
    \includegraphics[
      width=0.22\textwidth,
      trim=0 75 0 25,
      clip
    ]{assets/vogels.jpg}
  };
}

\newcommand{\hoare}{
  \node[pic, right=0.2cm of vogels] (hoare) {
    \includegraphics[width=0.22\textwidth]{assets/hoare.jpg}
  };
}

\newcommand{\lamport}{
  \node[pic, right=0.2cm of hoare] (lamport) {
    \includegraphics[
      width=0.22\textwidth,
      trim=250 700 300 200,
      clip
    ]{assets/lamport.jpg}
  };
}

\newcommand{\brewer}{
  \node[pic, right=0.2cm of lamport] (brewer) {
    \includegraphics[
      width=0.22\textwidth,
      trim=100 300 100 100,
      clip
    ]{assets/brewer.jpg}
  };
}

\begin{frame}
  \note{%
    Hi everyone! Thanks so much for attending. I'm Michael, and today I'll be
    talking about ``Interactive Checks for Coordination Avoidance''. This is
    joint work with my advisor at UC Berkeley, Joe Hellerstein.

    I'd like to begin my talk, if you don't mind, with a short fairy tale. The
    fairy tale involves an application developer who has some data they need to
    replicate. Unfortunately, the application developer cannot decide whether
    to replicate with weak consistency or strong consistency. The application
    developer stays up all night trying to decide but eventually falls asleep
    at their desk. In their sleep, they are visited by four wise men.

    % Imagine you're an application developer and you're trying to decide what
    % kind of consistency you'd like to use to replicate some of your data.

  }
\end{frame}

\begin{frame}
  \begin{center}
    \begin{tikzpicture}[xscale=1.3]
      \vogels
      \node[quote, text width=0.8\textwidth, align=left] at (0, 3) {
        \Large\textit{You should use weak consistency. It's wicked fast!}
      };
      \path[fill=gray!30] (vogels.north) -- ++(0, 1) -- ++(0.5, 0) -- (vogels.north);
    \end{tikzpicture}
  \end{center}

  \note{%
    % Someone like Werner Vogels would probably tell you to use weak consistency.
    % You can implement weak consistency super duper fast, and for many
    % applications, weak consistency is good enough.

    ``Weak. Weak. It's weak that you're seeking'', squeaked Werner Vogels.
    His voice trembled, creaking, as he kept speaking.
    ``Please learn that if you yearn your performance to peak,
    then turn no further than weak.'' critiqued Werner.

    Werner Vogels has a good point. You can implement weak consistency super
    duper fast, and for many applications, weak consistency is good enough. But
    anyway, let's continue.
  }
\end{frame}

\begin{frame}
  \begin{center}
    \begin{tikzpicture}[xscale=1.3]
      \vogels
      \hoare
      \node[quote, text width=0.8\textwidth, align=left] at (1, 3) {
        \Large\textit{What about the application invariants?!}
      };
      \path[fill=gray!30] (hoare.north) -- ++(0, 1.5) -- ++(1, 0) -- (hoare.north);
    \end{tikzpicture}
  \end{center}

  \note{%
    % But then Tony Hoare laments, ``no no no''. Hoare encourages us not to
    % settle for weak consistency because a weakly consistent system is hard to
    % reason about. It robs us of a fundamental way we use to reason about
    % programs: application invariants. Even if every transaction in a weakly
    % consistent system preserves application invariants, the whole system may
    % not.

    ``No, no, no more'' groaned Tony Hoare.
    He moaned, he lamented, his soul was tormented
    ``Weak is not easy. Weak is just treason.
    Please, these anomalies, make it so hard to reason.''

    You see, Hoare encourages us not to settle for weak consistency because a
    weakly consistent system is very hard to reason about.
  }
\end{frame}

\begin{frame}
  \begin{center}
    \begin{tikzpicture}[xscale=1.3]
      \vogels
      \hoare
      \lamport
      \node[quote, text width=0.8\textwidth, align=left] at (2, 3) {
        \Large\textit{Yeah, go with strong consistency! I have an algorithm for that :)}
      };
      \path[fill=gray!30] (lamport.north) -- ++(0, 1.5) -- ++(1, 0) -- (lamport.north);
    \end{tikzpicture}
  \end{center}

  \note{%
    % Leslie Lamport chimes in agreeingly. He encourages us to choose strong
    % consistency. Strongly consistent systems are much easier to reason about.
    % After all a strongly consistent system to end users looks exactly like a
    % system that's not even distributed.

    ``You see, I agree'', says he, the ever mesmerizing Leslie.
    ``Weak is just wrong, not right for long.
    To defend your end users, you must go with strong.''

    Leslie Lamport encourages us to choose strong consistency. Strongly
    consistent systems are much easier to reason about. After all a strongly
    consistent system to end users looks exactly like a system that's not even
    distributed.
  }
\end{frame}

\begin{frame}
  \begin{center}
    \begin{tikzpicture}[xscale=1.3]
      \vogels
      \hoare
      \lamport
      \brewer
      \node[quote, text width=0.8\textwidth, align=left] at (3, 3) {
        \Large\textit{Not so fast!}
      };
      \path[fill=gray!30] (brewer.north) -- ++(0, 1.5) -- ++(-1, 0) -- (brewer.north);
    \end{tikzpicture}
  \end{center}

  \note{%
    % ``Aha, not so fast'' counters Eric Brewer! Consistency and availability and
    % partition tolerance, oh my. Surely, strong consistency comes at the price
    % of availability and performance.
    %
    % So we're in a bit of a pickle. Weak consistency is too confusing and strong
    % consistency is too heavy handed.

    ``Alas, not so fast'', Brewer the dapper chap clapped back.
    ``Consistency weak but not weary, strong but not nearly
    as fast as you think when you know my CAP theory.''

    Eric Brewer is right. Surely, strong consistency comes at the price of
    availability and performance. So our application developer is in a bit of a
    pickle. Weak consistency is too confusing and strong consistency is too
    heavy handed. Fortunately, a hero emerges to save the day.
  }
\end{frame}

\begin{frame}
  \begin{center}
    \begin{tikzpicture}[xscale=1.3]
      \node[pic] (bailis) at (0, 0) {\includegraphics[width=0.25\textwidth]{assets/bailis.jpg}};
      \node[quote, text width=0.8\textwidth, align=left] at (0, 3) {
        \Large\textit{?`Porque no los dos?}
      };
      \path[fill=gray!30] (bailis.north) -- ++(0, 1.5) -- ++(1, 0) -- (bailis.north);
    \end{tikzpicture}
  \end{center}

  \note{%
    But then out teetered Peter, with swagger demeanor.
    ``Why not have both?'' quoth Peter.
    The end.

    Peter Bailis comes to save the day and reminds us that perhaps we can have
    the best of both worlds. In 2014, Peter and some other smart folks
    including Alan defined invariant confluence as a condition under which we
    can replicate a piece of data without the need for strongly consistent
    coordination while maintaining the data's application invariants.

    Invariant confluence will be the topic of this talk. In particular, I'll
    define invariant confluence and then discuss how we can check invariant
    confluence automatically.
  }
\end{frame}