### 2019-08-16 doc/vldb_poster/sections/iconfluence.tex
Finished poster.

\begin{itemize}
  \item
    A \defword{distributed object} $O = (S, \join)$ consists of a set $S$ of
    states and a binary merge operator $\join: S \times S \to S$.
  \item
    A \defword{transaction} $t: S \to S$ is a function.
  \item
    An \defword{invariant} $I$ is a subset of $S$.
\end{itemize}

\tikzstyle{s0color}=[fill=flatred]
\tikzstyle{s1color}=[fill=flatgreen]
\tikzstyle{s2color}=[fill=flatdenim]
\tikzstyle{s3color}=[fill=flatorange]
\tikzstyle{s4color}=[fill=flatyellow]
\tikzstyle{s5color}=[fill=flatcyan]
\tikzstyle{s6color}=[fill=flatpurple]
\tikzstyle{s7color}=[fill=flatblue]
\tikzstyle{txnline}=[black, line width=6pt, -latex]

\newcommand{\internaltext}[1]{$\boldsymbol #1$}

\tikzstyle{phantomstate}=[%
  shape=circle, inner sep=12pt, draw=white, line width=4pt, fill=white]
\tikzstyle{state}=[%
  shape=circle, inner sep=12pt, draw=black, line width=4pt, text opacity=1,
  fill opacity=0.6]

\begin{center}
  \large
  \begin{tikzpicture}[scale=5]
    \node (p3) at (0.5, 3) {$p_1$};
    \node (p2) at (0.5, 2) {$p_2$};
    \node (p1) at (0.5, 1) {$p_3$};

    \tikzstyle{pline}=[gray, opacity=0.75, line width=6pt, -latex]
    \draw[pline] (p3) to (4.75, 3);
    \draw[pline] (p2) to (4.75, 2);
    \draw[pline] (p1) to (4.75, 1);

    % Top line.
    \node[phantomstate] (s03) at (1, 3) {$\phantom{s_0}$};
    \node[phantomstate] (s13) at (2, 3) {$\phantom{s_1}$};
    \node[phantomstate] (s23) at (3, 3) {$\phantom{s_3}$};
    \node[phantomstate] (s33) at (3.75, 3) {$\phantom{s_6}$};
    \node[state, s0color] (s03) at (1, 3) {\internaltext{s_0}};
    \node[state, s1color] (s13) at (2, 3) {\internaltext{s_1}};
    \node[state, s3color] (s23) at (3, 3) {\internaltext{s_3}};
    \node[state, s6color] (s33) at (3.75, 3) {\internaltext{s_6}};

    % Middle line.
    \node[phantomstate] (s02) at (1, 2) {$\phantom{s_0}$};
    \node[phantomstate] (s12) at (2, 2) {$\phantom{s_2}$};
    \node[phantomstate] (s22) at (3, 2) {$\phantom{s_4}$};
    \node[state, s0color] (s02) at (1, 2) {\internaltext{s_0}};
    \node[state, s2color] (s12) at (2, 2) {\internaltext{s_2}};
    \node[state, s4color] (s22) at (3, 2) {\internaltext{s_4}};

    % Bottom line.
    \node[phantomstate] (s01) at (1, 1) {$\phantom{s_0}$};
    \node[phantomstate] (s11) at (2, 1) {$\phantom{s_2}$};
    \node[phantomstate] (s21) at (3, 1) {$\phantom{s_5}$};
    \node[phantomstate] (s31) at (4.25, 1) {$\phantom{s_7}$};
    \node[state, s0color] (s01) at (1, 1) {\internaltext{s_0}};
    \node[state, s2color] (s11) at (2, 1) {\internaltext{s_2}};
    \node[state, s5color] (s21) at (3, 1) {\internaltext{s_5}};
    \node[state, s7color] (s31) at (4.25, 1) {\internaltext{s_7}};

    \tikzstyle{txntext}=[sloped, above]
    \draw[txnline] (s03) to node[txntext]{$t$} (s13);
    \draw[txnline] (s13) to node[txntext]{$v$} (s23);
    \draw[txnline] (s02) to node[txntext]{$u$} (s12);
    \draw[txnline] (s01) to node[txntext]{$u$} (s11);
    \draw[txnline] (s11) to node[txntext]{$w$} (s21);

    \draw[txnline] (s13) to node[txntext]{} (s22);
    \draw[txnline] (s12) to node[txntext]{} (s22);
    \draw[txnline] (s22) to node[txntext]{} (s33);
    \draw[txnline] (s23) to node[txntext]{} (s33);
    \draw[txnline] (s21) to node[txntext]{} (s31);
    \draw[txnline] (s33) to node[txntext]{} (s31);
  \end{tikzpicture}
\end{center}

\[
  e ::= s \mid t(e) \mid e_1 \join e_2
\]

\begin{center}
  \begin{tikzpicture}[scale=4]
    \node[state, s7color, label={[label distance=-0.1cm] 90:$s_7$}] (j1) at (0, 0) {\internaltext{\join}};
    \draw (j1)++(-30:1) node[state, s5color, label={[label distance=-0.1cm] 90:$s_5$}] (w)            {\internaltext{w}};
    \draw (j1)++(210:1) node[state, s6color, label={[label distance=-0.1cm] 90:$s_6$}] (j2)           {\internaltext{\join}};
    \draw (w)++(-90:1)  node[state, s2color, label={[label distance=-0.2cm] 60:$s_2$}] (u1)           {\internaltext{u}};
    \draw (u1)++(-90:1) node[state, s0color]                                           (s4)           {\internaltext{s_0}};
    \draw (j2)++(225:1) node[state, s3color, label={[label distance=-0.1cm] 90:$s_3$}] (v)            {\internaltext{v}};
    \draw (v)++(-90:1)  node[state, s1color, label={[label distance=-0.2cm]120:$s_1$}] (t2)           {\internaltext{t}};
    \draw (t2)++(-90:1) node[state, s0color]                                           (s1)           {\internaltext{s_0}};
    \draw (j2)++(-45:1) node[state, s4color, label={[label distance=-0.1cm] 90:$s_4$}] (j3)           {\internaltext{\join}};
    \draw (j3)++(240:1) node[state, s1color, label={[label distance=-0.2cm]120:$s_1$}] (t3)           {\internaltext{t}};
    \draw (t3)++(-90:1) node[state, s0color]                                           (s2)           {\internaltext{s_0}};
    \draw (j3)++(-60:1) node[state, s2color, label={[label distance=-0.1cm] 90:$s_2$}] (u3)           {\internaltext{u}};
    \draw (u3)++(-90:1) node[state, s0color]                                           (s3)           {\internaltext{s_0}};

    \tikzstyle{astedge}=[line width=6pt]
    \draw[astedge] (j1) to (w) to (u1) to (s4);
    \draw[astedge] (j1) to (j2) to (v) to (t2) to (s1);
    \draw[astedge] (j1) to (j2) to (j3) to (t3) to (s2);
    \draw[astedge] (j1) to (j2) to (j3) to (u3) to (s3);
  \end{tikzpicture}
\end{center}

\begin{mathpar}
  \inferrule{ }{\sTIreachablepredicate{s_0}}

  \inferrule{\sTIreachablepredicate{e} \\ I(t(e))}
            {\sTIreachablepredicate{t(e)}}

  \inferrule{\sTIreachablepredicate{e_1} \\ \sTIreachablepredicate{e_2}}
            {\sTIreachablepredicate{e_1 \join e_2}}
\end{mathpar}

$O$ is \defword{\invariantconfluent} with respect to $s_0$, $T$, and $I$,
abbreviated \defword{\sTIconfluent{}}, if all reachable states satisfy the
invariant:
\[
  \setst{s \in S}{\sTIreachablepredicate{s}} \subseteq I
\]