### 2020-01-19 doc/the_vldb_journal_paper/figures/eval_table.tex
Formatted captions for Journal requirements.
They don't want punctuation at the end of a caption, and they want table
captions to describe the contents.

\begin{table}[t]
  \centering
  \caption{%
    \exampleref{TwoIntsEval} to \exampleref{TpccEval} summary. For every
    example, we list the time required (in seconds) to run our decision
    procedures to completion (excluding human input time) as well as the number
    of lines of code to specify the example
  }%
  \tablabel{EvalTable}
  \begin{tabular}{lrr}
    \toprule
    Example                                          & Run time (s) & Lines of code \\\midrule
    \ref{exa:TwoIntsEval}                            & 0.09         & 7 \\
    \ref{exa:ForeignKeysEval} (all transactions)     & 0.06         & 8 \\
    \ref{exa:ForeignKeysEval} (limited transactions) & 0.09         & 10 \\
    \ref{exa:AuctionEval}                            & 0.04         & 21 \\
    \ref{exa:EscrowTransactionsEval}                 & 0.09         & 49 \\
    \ref{exa:TpccEval} (Invariant 1)                 & 0.46         & 66 \\
    \ref{exa:TpccEval} (Invariant 2)                 & 0.44         & 33 \\
    \bottomrule
  \end{tabular}
\end{table}