### 2020-05-16 doc/sigmod_paper/invariant_confluence.tex
Copied VLDB to SIGMOD.
I copied over the VLDB paper but ported it to the SIGMOD style files.
Now, we just have to trim!

\documentclass{research4cacm}

\usepackage[noend]{algpseudocode}
\usepackage{algorithm}
\usepackage{balance}
\usepackage{booktabs}
\usepackage{invariantconfluence}
\usepackage{listings}
\usepackage{pervasives}
\usepackage{printlen}
\usepackage{python}
\usepackage{subcaption}
\usepackage{tikz}
\usetikzlibrary{arrows}
\usetikzlibrary{calc}
\usetikzlibrary{patterns}

\begin{document}
\title{%
  TODO%
  \thanks{%
    The original version of this paper is entitled ``Interactive Checks for
    Coordination Avoidance'' and was published in PVLDB Vol.\ 12.
  }%
}

\numberofauthors{2}
\author{
\alignauthor
Michael Whittaker\\
       \affaddr{UC Berkeley}\\
       \affaddr{Berkeley, CA}\\
       \email{mjwhittaker@berkeley.edu}
\alignauthor
Joseph M. Hellerstein\\
       \affaddr{UC Berkeley}\\
       \affaddr{Berkeley, CA}\\
       \email{hellerstein@berkeley.edu}
}

\maketitle

{\input{sections/abstract}}
{\input{sections/introduction}}
{\input{sections/invariant_confluence}}
{\input{sections/invariant_closure}}
{\input{sections/interactive_decision_procedure}}
{\input{sections/merge_reduction}}
{\input{sections/segmented_invariant_confluence.tex}}
{\input{sections/evaluation}}
{\input{sections/related_work}}
{\input{sections/conclusion}}

%ACKNOWLEDGMENTS are optional
\section{Acknowledgments}
This section is optional; it is a location for you
to acknowledge grants, funding, editing assistance, collaboration, access to
uncommon, but necessary, hardware or software etc.
In the present case, for example, the
authors would like to thank Gerald Murray of ACM for
his help in codifying this \textit{Author's Guide}
and the \textbf{.cls} and \textbf{.tex} files that it describes.

\bibliographystyle{abbrv} % standard abbrv style
\bibliography{references}
\balancecolumns
\end{document}