### 2020-05-16 doc/sigmod_paper/invariant_confluence.tex
Cut merge reduction section; split acks out.

\documentclass{research4cacm}

\usepackage[noend]{algpseudocode}
\usepackage{algorithm}
\usepackage{balance}
\usepackage{booktabs}
\usepackage{invariantconfluence}
\usepackage{listings}
\usepackage{pervasives}
\usepackage{printlen}
\usepackage{python}
\usepackage{subcaption}
\usepackage{tikz}
\usetikzlibrary{arrows}
\usetikzlibrary{calc}
\usetikzlibrary{patterns}

\begin{document}
\title{%
  TODO%
  \thanks{%
    The original version of this paper is entitled ``Interactive Checks for
    Coordination Avoidance'' and was published in PVLDB Vol.\ 12.
  }%
}

\numberofauthors{2}
\author{
\alignauthor
Michael Whittaker\\
       \affaddr{UC Berkeley}\\
       \affaddr{Berkeley, CA}\\
       \email{mjwhittaker@berkeley.edu}
\alignauthor
Joseph M. Hellerstein\\
       \affaddr{UC Berkeley}\\
       \affaddr{Berkeley, CA}\\
       \email{hellerstein@berkeley.edu}
}

\maketitle

{\input{sections/abstract}}
{\input{sections/introduction}}
{\input{sections/invariant_confluence}}
{\input{sections/invariant_closure}}
{\input{sections/interactive_decision_procedure}}
{\input{sections/segmented_invariant_confluence.tex}}
{\input{sections/evaluation}}
{\input{sections/related_work}}
{\input{sections/conclusion}}
{\input{sections/acknowledgements}}

\bibliographystyle{abbrv} % standard abbrv style
\bibliography{references}
\balancecolumns
\end{document}