### 2020-05-16 doc/sigmod_paper/sections/abstract.tex
Fixed copyright.
The publisher gave us some specific instructions on what the copyright
should say. Hopefully, I've done this right.

\begin{abstract}
Strongly consistent distributed systems are easy to reason about but face
fundamental limitations in availability and performance. Weakly consistent
systems can be implemented with very high performance but place a burden on the
application developer to reason about complex interleavings of execution.
\Invariantconfluence{} provides a formal framework for understanding when we
can get the best of both worlds. An \invariantconfluent{} object can be
efficiently replicated with no coordination needed to preserve its invariants.
However, actually determining whether or not an object is \invariantconfluent{}
is challenging.

In this paper, we establish conditions under which a commonly used sufficient
condition for \invariantconfluence{} is both necessary and sufficient, and we
use this condition to design a general-purpose interactive
\invariantconfluence{} decision procedure. We then take a step beyond
\invariantconfluence{} and introduce a generalization of
\invariantconfluence{}, called segmented \invariantconfluence{}, that allows us
to replicate non-\invariantconfluent{} objects with a small amount of
coordination.
%
We implemented these formalisms in a prototype called Lucy and found that our
decision procedures efficiently handle common real-world workloads including
%
% foreign keys, rollups, escrow transactions, and more.
foreign keys, escrow transactions, and more.
%
We also found that segmented \invariantconfluent{} replication can deliver up
to an order of magnitude more throughput than linearizable replication.
% for low contention workloads and comparable throughput for medium to high
% contention workloads.
\end{abstract}