### 2020-05-16 doc/sigmod_paper/sections/conclusion.tex
First round of cuts.
The paper is now 8 pages without citations. We need 8 pages with
citations, so we need to trim another page or so. I'll probably shrink
the evaluation a bit more, shrink the related works a bit more, shrink
segmented invariant confluence a bit more and see what we end up with.

\section{Conclusion}
This paper revolved around two major contributions.
%
First,
% we developed a deeper understanding of \invariantclosure{} and
% \invariantconfluence{} by looking at the two criteria with reachability in
% mind.
we found that \invariantclosure{} fails to incorporate a notion of
reachability, and using this intuition, we developed conditions under which
\invariantclosure{} and \invariantconfluence{} are equivalent. We implemented
this insight in an interactive \invariantconfluence{} decision procedure
that automatically checks whether an object is \invariantconfluent{}, with the
assistance of a programmer.
%
Second, we proposed a new consistency model and generalization of
\invariantconfluence{}, segmented \invariantconfluence{}, that can be used to
replicate non-\invariantconfluent{} objects with a small amount of coordination
while still preserving their invariants.
% We found that segmented
% \invariantconfluence{} naturally subsumes existing techniques for maintaining
% invariants of replicated objects (e.g.\ locking and escrow transactions), and
% we developed an interactive decision procedure for segmented
% \invariantconfluence{}.

% Through our evaluation, we found that our decision procedures could analyze a
% number of realistic workloads, each in less than a second. We also showed that
% segmented \invariantconfluence{} can significantly outperform linearizable
% replication for low-coordination workloads.