### 2020-05-16 doc/sigmod_paper/sections/introduction.tex

\section{Introduction}
When an application designer decides to replicate a piece of data, they have to
make a fundamental choice between weak and strong consistency. Replicating the
data with strong consistency using a technique like distributed transactions
(e.g., \cite{bernstein1981concurrency,mohan1986transaction}) or state machine
replication (e.g., \cite{schneider1990implementing, lamport1998part,
liskov2012viewstamped, ongaro2014search}) makes the application designer's life
very easy. To the developer, a strongly consistent system behaves exactly like
a single-threaded system running on a single node, so reasoning about the
behavior of the system is simple~\cite{herlihy1990linearizability}.
Unfortunately, strong consistency is at odds with performance. The CAP theorem
and PACELC theorem tell us that strongly consistent systems suffer from higher
latency at best and unavailability at worst~\cite{gilbert2002brewer,
brewer2012cap, abadi2012consistency}. On the other hand, weak consistency
models like eventual consistency~\cite{vogels2009eventually}, PRAM
consistency~\cite{lipton1988pram}, causal consistency~\cite{ahamad1995causal},
and others~\cite{lloyd2011don, mehdi2017can} allow data to be replicated with
high availability and low latency, but they put a tremendous burden on the
application designer to reason about the complex interleavings of operations
that are allowed by these weak consistency models. In particular, weak
consistency models strip an application developer of one of the earliest and
most effective tools that is used to reason about the execution of programs:
application invariants~\cite{hoare1969axiomatic, balegas2015towards} such as
database integrity constraints~\cite{grefen1993integrity, gupta1993local}. Even
if every transaction executing in a weakly consistent system individually
maintains an application invariant, the system as a whole can produce
invariant-violating states.

Is it possible for us to have our strongly consistent cake and eat it with high
availability too? Can we replicate a piece of data with weak consistency but
still ensure that its invariants are maintained? Yes... sometimes. Bailis et
al.\ introduced the notion of \emph{\invariantconfluence{}} as a necessary and
sufficient condition for when invariants can be maintained over replicated data
without the need for any coordination~\cite{bailis2014coordination}. If an
object is \invariantconfluent{} with respect to an invariant, we can replicate
it with the performance benefits of weak consistency and (some of) the
correctness benefits of strong consistency.

Unfortunately, to date, the task of identifying whether or not an object
actually is \invariantconfluent{} has remained an exercise in human proof
generation. Bailis et al.\ manually categorized a set of common objects,
transactions, and invariants (e.g.\ foreign key constraints on relations,
linear constraints on integers) as \invariantconfluent{} or not. Hand-written
proofs of this sort are unreasonable to expect from programmers. Ideally we
would have a general-purpose program that can automatically determine
\invariantconfluence{} for us.  \textbf{The first main thrust of this paper is
to make \invariantconfluence{} checkable:} to design a general-purpose
\invariantconfluence{} decision procedure, and implement it in an interactive
system.

Unfortunately, designing such a general-purpose decision procedure is
impossible because determining the \invariantconfluence{} of an object is
undecidable in general. Still, we can develop a decision procedure that works
well in the common case.
%
For example, many prior efforts have developed decision procedures for
\emph{\invariantclosure{}}, a sufficient (but not necessary) condition for
\invariantconfluence{}~\cite{li2012making, li2014automating}. The existing
approaches check whether an object is \invariantclosed{}. If it is, then they
conclude that it is also \invariantconfluent{}. If it's not, then the current
approaches are unable to conclude anything about whether or not the object is
\invariantconfluent{}.

In this paper, we take a step back and study the underlying reason \emph{why}
\invariantclosure{} is not necessary for \invariantconfluence{}.
%
% We find that \invariantconfluence{} is fundamentally a reachability property,
% checking whether certain values of our replicated object can actually be
% obtained in the execution of a system. We discover that unreachable, yet
% invariant satisfying, states cause \invariantclosure{} to be an unnecessary
% condition.
%
Using this understanding, we construct a set of modest conditions under which
\invariantclosure{} and \invariantconfluence{} are in fact \emph{equivalent},
allowing us to reduce the problem of determining \invariantconfluence{} to that
of determining \invariantclosure{}.
%
Then, we use these conditions to design a general-purpose interactive
\invariantconfluence{} decision procedure.

% Then, we use these conditions to design a general-purpose interactive
% \invariantconfluence{} decision procedure. Users iteratively interact with the
% decision procedure, eliminating unreachable states from the invariant, moving
% towards the conditions under which \invariantclosure{} and
% \invariantconfluence{} are equivalent. Meanwhile, the decision procedure
% generates examples of reachable and unreachable states which help the user
% recognize patterns that describe reachability. We also leverage our intuitions
% about reachability to develop a new sufficient condition for
% \invariantconfluence{}, dubbed \emph{\mergereducibility}. \Mergereducibility{}
% covers some cases that are not covered by \invariantclosure{}, and it can be
% automatically checked without any user interaction.

\textbf{The second main thrust of this paper is to partially avoid coordination
even in programs that require it}, by generalizing \invariantconfluence{} to a
property called \emph{segmented \invariantconfluence{}}. While
\invariantconfluence{} characterizes objects that can be replicated
\emph{without any} coordination, segmented \invariantconfluence{} allows us to
replicate non-\invariantconfluent{} objects with only \emph{occasional}
coordination. The main idea is to divide the set of invariant-satisfying states
into \emph{segments}, with a restricted set of transactions allowed in each
segment.  Within a segment servers act without any coordination; they
synchronize only to transition across segment boundaries. This design
highlights the trade-off between application complexity and
coordination-freedom; more complex applications require more segments which
require more coordination, and vice-versa.

Finally, we present Lucy: an implementation of our decision procedures and a
system for replicating \invariantconfluent{} and segmented
\invariantconfluent{} objects. Using Lucy, we find that our decision procedures
can efficiently handle a wide range of common workloads.  For example, in
\secref{Evaluation}, we apply Lucy to foreign key constraints and escrow
transactions. Lucy processes these workloads in less than half a second, and
neither workload requires more than 66 lines of code to specify. Moreover, we
find that segmented \invariantconfluent{} replication can achieve 10x to 100x
more throughput than linearizable replication for low-coordination workloads.

% In closing, here is an outline of the paper and of our contributions:
% %
% We propose a novel expression-oriented definition of \invariantconfluence{}
% that is both formal and simple (\secref{InvariantConfluence}).
% %
% We develop an understanding of why \invariantclosure{} is not necessary for
% \invariantconfluence{} and use this understanding to develop conditions under
% which it is both necessary and sufficient (\secref{InvariantClosure}).
% %
% We exploit these conditions to design a general-purpose interactive decision
% procedure for \invariantconfluence{} (\secref{InteractiveDecisionProcedure}).
% %
% We again exploit these conditions to design a novel non-trivial sufficient
% condition for \invariantconfluence{}, called \mergereducibility.
% %
% We present segmented \invariantconfluence{}: a generalization of
% \invariantconfluence{} that uses a small amount of coordination to maintain
% invariants for replicated objects that are otherwise not \invariantconfluent{}
% (\secref{SegmentedInvariantConfluence}).
% %
% We evaluate our methods using a prototype implementation called Lucy
% (\secref{Evaluation}).