### 2020-05-22 doc/sigmod_paper/sections/conclusion.tex

\section{Conclusion}
This paper revolved around two major contributions. First, we found that
\invariantclosure{} fails to incorporate a notion of reachability, and using
this intuition, we developed conditions under which \invariantclosure{} and
\invariantconfluence{} are equivalent. We implemented this insight in an
interactive \invariantconfluence{} decision procedure that automatically checks
whether an object is \invariantconfluent{}, with the assistance of a
programmer.
%
Second, we proposed a new consistency model and generalization of
\invariantconfluence{}, segmented \invariantconfluence{}, that can be used to
replicate non-\invariantconfluent{} objects with a small amount of coordination
while still preserving their invariants.