### 2020-06-14 doc/sigmod_paper/sections/invariant_confluence.tex

\section{Invariant Confluence}\seclabel{InvariantConfluence}
Informally, a replicated object is \defword{\invariantconfluent{}} with respect
to an invariant if every replica of the object is guaranteed to satisfy the
invariant despite the possibility of different replicas being concurrently
modified or merged together~\cite{bailis2014coordination}. In this section, we
make this informal notion of \invariantconfluence{} precise.

We begin by introducing our system model of replicated objects in which a
distributed object and an invariant are replicated across a set of
servers. Clients send transactions to servers, and servers execute
transactions so long as they maintain the invariant. Servers execute
transactions without coordination, but to avoid state divergence, servers
periodically gossip with one another and merge their replicas.
%
% After we introduce the system model, we present a formal definition of
% \invariantconfluence{}.

\subsection{System Model}

A \defword{distributed object} $O = (S, \join)$ consists of a set $S$ of states
and a binary merge operator $\join: S \times S \to S$ that merges two states
into one. A \defword{transaction} $t: S \to S$ is a function that maps one
state to another. An \defword{invariant} $I$ is a subset of $S$. Notationally,
we write $I(s)$ to denote that $s$ satisfies the invariant (i.e. $s \in I$)
and $\lnot I(s)$ to denote that $s$ does not satisfy the invariant (i.e. $s
\notin I$).

\vfill\null

\begin{example}\examplelabel{Ints}
  $O = (\ints, \max)$ is a distributed object consisting of integers merged by the
  $\max$ function; $t(x) = x + 1$ is a transaction that adds one to a state; and
  $\setst{x \in \ints}{x \geq 0}$ is the invariant that states $x$ are
  non-negative.
\end{example}

In our system model, a distributed object $O$ is replicated across a set $p_1,
\ldots, p_n$ of $n$ servers. Each server $p_i$ manages a replica $s_i \in S$ of
the object. Every server begins with a start state $s_0 \in S$, a
fixed set $T$ of transactions, and an invariant $I$. Servers repeatedly perform
one of two actions.

First, a client can contact a server $p_i$ and request that it execute a
transaction $t \in T$. $p_i$ speculatively executes $t$, transitioning from
state $s_i$ to state $t(s_i)$. If $t(s_i)$ satisfies the invariant---i.e.
$I(t(s_i))$---then $p_i$ commits the transaction and remains in state $t(s_i)$.
Otherwise, $p_i$ aborts the transaction and reverts to state $s_i$.

Second, a server $p_i$ can send its state $s_i$ to another server $p_j$ with
state $s_j$ causing $p_j$ to transition from state $s_j$ to state $s_i \join
s_j$. Servers periodically merge states with one another in order to keep their
states loosely synchronized. Note that unlike with transactions, servers
\emph{cannot} abort a merge; server $p_j$ must transition from $s_j$ to $s_i
\join s_j$ whether or not $s_i \join s_j$ satisfies the invariant.

Informally, $O$ is \defword{\invariantconfluent{}} with respect to $s_0$, $T$,
and $I$, abbreviated \defword{\sTIconfluent{}}, if every replica $s_1, \ldots,
s_n$ is guaranteed to always satisfy the invariant $I$ in every possible
execution of the system.

\subsection{Expression-Based Formalism}
To define \invariantconfluence{} formally, we represent a state produced by a
system execution as a simple expression generated by the grammar
\[
  e ::= s \mid t(e) \mid e_1 \join e_2
\]
where $s$ represents a state in $S$ and $t$ represents a transaction in $T$. As
an example, consider the system execution in \figref{SystemExecution} in which
a distributed object is replicated across servers $p_1$, $p_2$, and $p_3$.
Server $p_3$ begins with state $s_0$, transitions to state $s_2$ by executing
transaction $u$, transitions to state $s_5$ by executing transaction $w$, and
then transitions to state $s_7$ by merging with server $p_1$. Similarly, server
$p_1$ ends up with state $s_6$ after executing transactions $t$ and $v$ and
merging with server $p_2$. In \figref{Expression}, we see the abstract syntax
tree of the corresponding expression for state $s_7$.

{\input{figures/invariant_confluence_definitions}}

We say an expression $e$ is \defword{\sTIreachable{}} if it corresponds to a
valid execution of our system model. Formally, we define
$\sTIreachablepredicate{e}$ to be the predicate that satisfies the
following conditions:
\pagebreak
\begin{itemize}
  \item
    $\sTIreachablepredicate{s_0}$.
  \item
    For all expressions $e$ and for all transactions $t$ in the set $T$ of
    transactions, if $\sTIreachablepredicate{e}$ and $I(t(e))$, then
    $\sTIreachablepredicate{t(e)}$.
  \item
    For expressions $e_1$ and $e_2$, if $\sTIreachablepredicate{e_1}$ and
    $\sTIreachablepredicate{e_2}$, then $\sTIreachablepredicate{e_1 \join
    e_2}$.
\end{itemize}

Similarly, we say a state $s \in S$ is \sTIreachable{} if there exists an
\sTIreachable{} expression $e$ that evaluates to $s$. Returning to
\exampleref{Ints} with start state $s_0 = 42$, we see that all integers greater
than or equal to 42 (i.e.\ $\setst{x \in \ints}{x \geq 42}$) are
\sTIreachable{}, and all other integers are \sTIunreachable{}.

Finally, we say $O$ is \defword{\invariantconfluent} with respect to $s_0$,
$T$, and $I$, abbreviated \defword{\sTIconfluent{}}, if all reachable states
satisfy the invariant:
\[
  \setst{s \in S}{\sTIreachablepredicate{s}} \subseteq I
\]